---
title: "Diagnose Linearity in DGVMs"
author: "Huanyuan Zhang-Zheng and Benjamin Stocker"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r}
library(tidyverse)
library(here)
# remotes::install_github("geco-bern/rgeco")
#library(rgeco)

source(here("R/collect_gdf_bymodl_HZ_Trendyv11.R"))
source(here("R/nc_to_df.R"))
source(here("R/read_nc_onefile.R"))
source(here("R/regrid_df.R"))
# # create script from this vignette
# knitr::purl(
#   here("vignettes/DGVMLIN_trendy_v11.rmd"), 
#   output = here("analysis/dgmvlin.R"), 
#   documentation = 1
#   )

#setwd("~/E/RA/DGVM_density_plot/trendy_v9") #h
#setwd('F:/Oxford/Chapter_four/dgvmlin')
# If you only want to draw figures, only do this: load(here("data/gdf_Trendyv11_202406.rda")) 
```
## If you just want to reproduce our figures

Step 1: run library and source trunk
Step 2: At regrid step, read in "data/gdf.rds"
Step 3: Run **Model meta info** and **Plot relationships** then you are free to jump to other figures drawing sections

## Approach

Does (steady-state) SOC storage, simulated in DGVMs, scale with litter inputs? And do litter inputs scale with biomass? To investigate this, we can look at the following relationships:

- The relative enhancement in steady-state SOC vs. the relative enhancement in NPP. $C^\ast$ is the steady-state SOC. It is given by the initial SOC stocks in simulations (spun up to steady state). 
$$
\frac{\Delta C^\ast}{C^\ast} / \frac{\Delta NPP}{NPP_\text{init}}
$$
- The relative enhancement in steady-state SOC vs. the relative enhancement in biomass ($B$).
$$
\frac{\Delta C^\ast}{C^\ast} / \frac{\Delta B}{B_\text{init}}
$$
$\Delta$ is the difference between end and beginning of a transient simulation covering the historical period, where only CO2 changes (hence, soil turnover rates should not change in a typical 1st order kinetics-based SOC model).

Our hypothesis is that these relationships follow the 1:1 line.

$\Delta C^\ast$ cannot be extracted directly from simulations. But it can be estimated from the imbalance between soil respiration and litter input before a steady state is reached:
$$
\tau \approx \frac{C(t)}{\text{NPP}(t) - \Delta C(t)}
$$

Combine this with $\Delta C^\ast=\tau \; \text{NPP}$, to get:
$$
C^\ast = \frac{C(t)}{1 - \frac{\Delta C(t)}{\text{NPP}(t)}}
$$
These variables are extracted from TRENDY S1 simulations as follows:

- $C(t)$: soil C at the end of the simulation (mean over 2008-2017)
- $\Delta C(t)$: Annual change in soil C during the last decade (2008-2017) (`filn_cSoil_change` - MUST BE DIVIDED BY 10!)
- NPP$(t)$: Annual mean NPP during the last decade (2008-2017)

- $\Delta C^\ast$: $C^\ast$ (as defined above) minus soil C during the first decade of the simulation (spun up to steady state)

The same is done with biomass ($B$).


## Process TRENDY v7 files

(Adjust path for `filnams` to select v8 or v7)

From downloaded TRENDY outputs (here, simulation S1: only CO2 is changing), get global fields of: 

1. Multi-annual mean NPP at the beginning of the simulation
2. Multi-annual mean NPP at the end of the simulation
3. Multi-annual mean SOC at the beginning of the simulation
4. Multi-annual mean SOC at the end of the simulation
5. Multi-annual mean biomass at the beginning of the simulation
6. Multi-annual mean biomass at the end of the simulation
7. Change in SOC over a multi-annual period at the end of the simulation (for $\Delta C(t)$)

Simulations should cover 1700-2017 (318 years = time step in NetCDF outputs). But not all do.

- Starting in year 1700: CABLE-POP, CLM5.0, JSBACH, LPX-Bern, LPJ-GUESS, OCN, ORCHIDEE-CNP, SDGVM
- Starting in year 1701: CLASS-CTEM, ISAM, JULES, SURFEX
- Starting in year 1900: DLEM
- Starting in year 1901: ORCHIDEE
- Starting in year 1860: VISIT

Peculiarities of outputs by model prohibiting their processing:

- LPJ: No S1 outputs
- DLEM: failed NPP processing
- LPX-Bern: failed NPP processing
- VISIT: No NPP output available

No separate output for pools (wood, leaf, root)

- CLASS-CTEM
- DLEM
- JSBACH
- JULES
- OCN
- SDGVM (only wood missing, could be derived by `cdo -O sub SDGVM_S1_cVeg.nc SDGVM_S1_croot.nc SDGVM_S1_cLeaf.nc`. cWood is then assumed zero: `cdo mulc,0 SDGVM_S1_cLeaf.nc SDGVM_S1_cWood.nc`, and `cdo -O chname,cLeaf,cWood SDGVM_S1_cWood.nc SDGVM_S1_cWood2.nc`, but is not!)
- SURFEX
- VISIT

### Process files: multi-annual mean

Determine for which models we have all the files (cWood and cVeg) to calculate wood mass fractions and create system command.
```{r}
filnams <- read_csv(here("data-raw/filnams_trendy_v11_S1.csv" )) %>% 
  setNames(c("modl", paste0("filn_", names(.)[-1])))

modls <- filnams %>% 
  pull(modl)

# This is a huge list of nc files that we will read
df <- filnams %>% 
  
  ## filter and correct typo
  filter(modl %in% modls) %>%
  mutate(dir = paste0(here(),"")) %>%
  
  # Please make sure R here() grab an address pointing to the data folder containing your nc files
  # my here() has "F:/Oxford/Chapter_four/dgvmlin", under this address it has a data folder
  mutate_at(vars(starts_with("filn")), ~str_replace(., "\\.nc", "")) %>%
  mutate_at(vars(starts_with("filn")), ~str_replace(., "cleaf", "cLeaf")) %>%
  mutate_at(vars(starts_with("filn")), ~str_replace(., "croot", "cRoot")) %>%
  mutate_at(vars(starts_with("filn")), ~str_replace(., "csoil", "cSoil")) %>%
  mutate_at(vars(starts_with("filn")), ~str_replace(., "cveg", "cVeg")) %>%
  mutate_at(vars(starts_with("filn")), ~str_replace(., "cwood", "cWood")) %>%
  mutate_at(vars(starts_with("filn")), ~str_replace(., "croot", "cRoot")) %>%
  
  ## get starting year
  left_join(read_csv(here("data-raw/startyear_trendy_v11_S1.csv" )), by = "modl") %>% 
  rename(startyear_init = startyear, startyear_npp_init = startyear_npp) %>% 
  mutate(endyear_init = startyear_init + 9, endyear_npp_init = startyear_npp_init + 9) %>% 
  
  rowwise() %>% 
  mutate(filn_cVeg_init  = paste0(filn_cVeg, "_INIT_MEAN"),
         filn_cVeg_final = paste0(filn_cVeg, "_FINAL_MEAN"),

         filn_cLeaf_init  = paste0(filn_cLeaf, "_INIT_MEAN"),
         filn_cLeaf_final = paste0(filn_cLeaf, "_FINAL_MEAN"),

         filn_cRoot_init  = paste0(filn_cRoot, "_INIT_MEAN"),
         filn_cRoot_final = paste0(filn_cRoot, "_FINAL_MEAN"),

         filn_cWood_init  = paste0(filn_cWood, "_INIT_MEAN"),
         filn_cWood_final = paste0(filn_cWood, "_FINAL_MEAN"),
         
         filn_cSoil_init  = paste0(filn_cSoil, "_INIT_MEAN"),
         filn_cSoil_final = paste0(filn_cSoil, "_FINAL_MEAN"),
         
         filn_cSoil_change = paste0(filn_cSoil, "_CHANGE"),
         filn_cVeg_change  = paste0(filn_cVeg, "_CHANGE"),
         filn_cLeaf_change = paste0(filn_cLeaf, "_CHANGE"),
         filn_cWood_change = paste0(filn_cWood, "_CHANGE"),
         filn_cRoot_change = paste0(filn_cRoot, "_CHANGE"),
                  
         filn_npp_init  = paste0(filn_npp, "_INIT_MEAN"),
         filn_npp_final = paste0(filn_npp, "_FINAL_MEAN"),
         
         filn_gpp_init  = paste0(filn_gpp, "_INIT_MEAN"),
         filn_gpp_final = paste0(filn_gpp, "_FINAL_MEAN"),
          
         filn_rh_init  = paste0(filn_rh, "_INIT_MEAN"),
         filn_rh_final = paste0(filn_rh, "_FINAL_MEAN"),
         
         filn_cLitter_init  = paste0(filn_cLitter, "_INIT_MEAN"),
         filn_cLitter_final = paste0(filn_cLitter, "_FINAL_MEAN"),
         
         filn_fRootLitter_final = paste0(filn_fRootLitter, "_FINAL_MEAN"),
         filn_fWoodLitter_final = paste0(filn_fWoodLitter, "_FINAL_MEAN")
         )

# Note that the above codes will report the lack of data and fill them with NA. Don't worry. Our analysis does not need quite all variables.



```

## Collect data

... into tidy data frame

Please note that Cwood_change Croot_change fRootLitter and fWoodLitter are for sensitivity test only. Only Jules have these things. For other models, the codes will warn you that these files are missing

```{r collect_data}
gdf <- purrr::map(

  as.list(seq(nrow(df))),
  ~collect_gdf_bymodl(

    df$modl[.],
    df$dir[.],

    df$filn_cSoil_init[.],
    df$filn_cSoil_final[.],

    df$filn_cVeg_init[.],
    df$filn_cVeg_final[.],

    df$filn_cLeaf_init[.],
    df$filn_cLeaf_final[.],

    df$filn_cRoot_init[.],
    df$filn_cRoot_final[.],

    df$filn_cWood_init[.],
    df$filn_cWood_final[.],

    df$filn_npp_init[.],
    df$filn_npp_final[.],

    df$filn_cSoil_change[.],
    df$filn_cVeg_change[.],
    df$filn_cWood_change[.],
    df$filn_cRoot_change[.],

    
    df$filn_gpp_init[.],
    df$filn_gpp_final[.],
    
    df$filn_rh_init[.],
    df$filn_rh_final[.],
    
    df$filn_cLitter_init[.],
    df$filn_cLitter_final[.],
    
    df$filn_landCoverFrac[.],
    df$filn_fRootLitter_final[.],
    df$filn_fWoodLitter_final[.]
    
    )) %>% 
  
  bind_rows() %>%

  group_by(modl) %>%
  nest()
```

## Cleaning

Remove outputs for gridcells that have zero GPP or cVeg at the start or at the end.
```{r remove_nonvegetated}
remove_nonvegetated <- function(df){
  df %>% 
    mutate(testvar = gpp_init * gpp_final * cveg_init * cveg_final) %>% 
    mutate(across(
      .cols = matches("(_init|_final|_change)$"),
      .fns = ~ ifelse(testvar == 0, NA, .)
    ))
}

desert_filter <- function(df){
  threshold <- quantile(df$cveg_init, probs = 0.05, na.rm = TRUE)
  df %>% 
    mutate(across(
      .cols = "cveg_init",
      .fns = ~ ifelse(cveg_init < threshold, NA, .)
    ))
}

desert_filter_wide <- function(df){
  df %>% 
    mutate(across(
      .cols = -c("lon", "lat"),
      .fns = ~ ifelse(is.na(cveg_init), NA, .)
    ))
}


gdf <- gdf %>% 
  mutate(
    data = map(data, ~remove_nonvegetated(.)),
    data = map(data, ~desert_filter(.)) # so we deal with NA on the cveg_init column first
 #   data = map(data, ~desert_filter_wide(.))  # if cveg_init is NA, everything else changed to NA. We do this in two steps so that we can see the effect of this processing sperately. 
    )
```

Longitude transformation for CLM5.0, ISAM, and JULES.
```{r tranform_lon}
shift_lon <- function(df){
  df |> 
    mutate(lon = ifelse(lon > 180, lon - 360, lon))
}

gdf <- gdf |> 
  mutate(data = map2(.x = data,
    .y = modl,
    .f = ~ if (.y %in% c("CLM5.0", "ISAM", "JULES")) shift_lon(.x) else .x
  ))
```

Overwrite `landCoverFrac` info for CABLE-POP. Many `NA` in this file, anyway no land cover change, we replace them with 0.
```{r correct_landcoverfrac}
gdf <- gdf |> 
  mutate(data = map2(.x = data,
    .y = modl,
    .f = ~ if (.y  == "CABLE-POP") mutate(.x, landCoverFrac = 0) else .x
  ))
```

## Calculate relationships

```{r calc_relationships}
calc_steady_state <- function(c_t, dcdt, npp){
  # c_t * 1/(1 - dcdt/npp)
  c_t * npp / (npp - dcdt)
}

calc_steady_state_alternative <- function(c_t, npp, rh){
  c_t * npp / rh
}

get_steadystate <- function(df){

  df %>%

    mutate(
      csoil_change = csoil_change / 10,
      cveg_change = cveg_change / 10,
      cRoot_change = cRoot_change/10,
       cWood_change = cWood_change/10, # because the change was calculated as the difference in the last 10 years
      npp_root = cRoot_change + fRootLitter,
      npp_wood = cWood_change + fWoodLitter
      ) %>%

    # applying clean function
    mutate(
      csoil_star_final = calc_steady_state(
        c_t = csoil_final,
        dcdt = csoil_change,
        npp = npp_final
        ),
      cveg_star_final = calc_steady_state(
        c_t = cveg_final,
        dcdt = cveg_change,
        npp = npp_final
        ),
      croot_star_final = calc_steady_state(
        c_t = croot_final,
        dcdt = cRoot_change,
        npp = npp_root
        ),
       cwood_star_final = calc_steady_state(
        c_t = cwood_final,
        dcdt = cWood_change,
        npp = npp_wood
        )
      )
}

get_deltas <- function(df){
  df %>%
    mutate(cveg_ag_init = cleaf_init + cwood_init,
           cveg_ag_final = cleaf_final + cwood_final,
           csoil_final = csoil_final + cLitter_final,
           csoil_init = csoil_init + cLitter_init
           ) %>%
    mutate(dcveg = (cveg_final - cveg_init)/cveg_init,
           dcveg_ag = (cveg_ag_final - cveg_ag_init)/cveg_ag_init,
           dnpp = (npp_final - npp_init)/npp_init,
           dcsoil = (csoil_final - csoil_init)/csoil_init,
           dcsoil_star = (csoil_star_final - csoil_init)/csoil_init,
           dgpp = (gpp_final - gpp_init)/gpp_init,
           dcleaf= (cleaf_final - cleaf_init)/cleaf_init,
           dcveg_star = (cveg_star_final - cveg_init)/cveg_init,
           dcroot = (croot_final - croot_init)/croot_init,
           dcwood = (cwood_final - cwood_init)/cwood_init,
           dcroot_star = (croot_star_final - croot_init)/croot_init,
           dcwood_star = (cwood_star_final - cwood_init)/cwood_init
           ) %>%
    mutate(l_npp_gpp = dnpp / dgpp,
           l_croot_cveg = dcroot / dcveg,
           l_croot_cveg_star = dcroot / dcveg_star,
           l_cveg_star_npp = dcveg_star / dnpp,
           l_cleaf_cveg = dcleaf / dcveg,
           l_cleaf_cveg_star = dcleaf / dcveg_star,
           l_cwood_cveg = dcwood / dcveg,
           l_cleaf_croot = dcleaf / dcroot
           )
}

filter_no_landcoverchange <- function(df){
  df %>%
    filter(landCoverFrac == 0)
}

gdf <- gdf %>%
  mutate(
    data = purrr::map(data, ~get_steadystate(.)),
    ) %>%
  mutate(
    data = purrr::map(data, ~get_deltas(.))
    ) %>%
  mutate(
    data = purrr::map(data, ~filter_no_landcoverchange(.))
    )
```

## Regrid -- ENTER HERE FOR REPRODUCING FIGURES

Regrid all model outputs to a common 1x1 deg grid. Note: this takes quite long, therefore reading from file if available.
```{r compile gdf}
filnam <- here("data/gdf_20260429.rds")

if (!file.exists(filnam)){
  
  gdf <- gdf %>% 
    slice(1) |> 
    mutate(
      data_1x1deg = purrr::map(
        data, 
        ~regrid_df(
          ., 
          res = 1.0, 
          lon_start = -180,
          lon_end = 180,
          lat_start = -90,
          lat_end = 90
          )
        )
      )

  readr::write_rds(gdf, file = filnam)
  
} else {
  
  gdf <- readr::read_rds(file = filnam)
  
}
```

## Model meta info

```{r}
summary_table <- tibble(
  modl = c(
    'ALL',
    'VISIT-NIES',
    'CLASSIC',
    'IBIS',
    'LPJ',
    'VISIT',
    'ISBA-CTRIP',
    'ORCHIDEE',
    'CLM5.0',
    'DLEM',
    'ISAM',
    'JSBACH',
    'JULES',
    'LPX-Bern',
    'SDGVM',
    'CABLE-POP',
    'LPJ-GUESS'
  ),
  cn_coupled = c(
    NA,
    FALSE, # VISIT-NIES
    FALSE, # CLASSIC
    FALSE, # IBIS
    FALSE, # LPJ
    FALSE, # VISIT
    FALSE, # ISBA-CTRIP
    TRUE,  # ORCHIDEE
    TRUE,  # CLM5.0
    TRUE,  # DLEM
    TRUE,  # ISAM
    TRUE,  # JSBACH
    TRUE,  # JULES
    TRUE,  # LPX-Bern
    TRUE,  # SDGVM
    TRUE,  # CABLE-POP
    TRUE   # LPJ-GUESS
  ),
  veg_demo = c(
    NA,
    FALSE, # VISIT-NIES
    FALSE, # CLASSIC
    FALSE, # IBIS
    FALSE, # LPJ
    FALSE, # VISIT
    FALSE, # ISBA-CTRIP
    FALSE, # ORCHIDEE
    FALSE, # CLM5.0
    FALSE, # DLEM
    FALSE, # ISAM
    FALSE, # JSBACH
    FALSE, # JULES
    FALSE, # LPX-Bern
    TRUE,  # SDGVM
    TRUE,  # CABLE-POP
    TRUE   # LPJ-GUESS
  )
)

# gdf <- gdf %>% 
#   left_join(
#     summary_table,
#     by = join_by(modl)
#   )

#readr::write_rds(gdf, file = here('data/gdf_before_regrid.rds'))
```

## Plot relationships

Preparations.
```{r}
get_p_value <- function(x, na.rm=TRUE) {
  p.value <- t.test(x, mu = 1)$p.value
  p.value
}

# sample for even representation, number determined by model with smallest 
# number of gridcells
# XXX this seems to have been updated to using more values per model. Please correct.
# nsample <- gdf %>% 
#   mutate(len = purrr::map_int(data, ~nrow(.))) %>% 
#   pull(len) %>% 
#   min()

# xxx adopted specification from paper here 
nsample <- 58697

coast <- rnaturalearth::ne_coastline(
  scale = 110, 
  returnclass = "sf"
  )

# 1. Define your custom order
model_order <- c(
  "VISIT-NIES", "CLASSIC", "IBIS", "LPJ", "VISIT", 
  "ISBA-CTRIP", "ORCHIDEE", "CLM5.0", "DLEM", "ISAM", 
  "JSBACH", "JULES", "LPX-Bern", "SDGVM", "CABLE-POP", "LPJ-GUESS"
)

display_titles <- c(
  "VISIT-NIES" = "VISIT-NIES",
  "CLASSIC"    = "CLASSIC",
  "IBIS"       = "IBIS",
  "LPJ"        = "LPJ",
  "VISIT"      = "VISIT",
  "ISBA-CTRIP" = "ISBA-CTRIP",
  "ORCHIDEE"   = "ORCHIDEE (C-N)",
  "CLM5.0"     = "CLM5.0 (C-N)",
  "DLEM"       = "DLEM (C-N)",
  "ISAM"       = "ISAM (C-N)",
  "JSBACH"     = "JSBACH (C-N)",
  "JULES"      = "JULES (C-N)",
  "LPX-Bern"   = "LPX-Bern (C-N)",
  "SDGVM"      = "SDGVM (C-N, Veg.D.)",
  "CABLE-POP"  = "CABLE-POP (C-N, Veg.D.)",
  "LPJ-GUESS"  = "LPJ-GUESS (C-N, Veg.D.)"
)

# 2. Ensure gdf$modl is a factor with this specific order

gdf$modl <- factor(gdf$modl, levels = model_order)
gdf <- gdf %>% arrange(modl)

model_order_with_all <- c("ALL",model_order)
display_titles_with_all <- c("ALL"="ALL",display_titles)
```


### GPP vs NPP 

#### All models pooled

```{r}
set.seed(1982)

gdf_sampled <- gdf %>%
  mutate(data = purrr::map(data, ~slice_sample(., n = nsample))) %>%
  unnest(data) %>%
  dplyr::filter(
    !is.nan(dnpp), !is.nan(dgpp), 
    !is.infinite(dnpp), !is.infinite(dgpp),
    !is.nan(l_npp_gpp), !is.infinite(l_npp_gpp))

# density scatter-plot for pooled data, balanced samples, original grid
gg_densityscatter <- gdf_sampled %>%
  ggplot(aes(x = dgpp, y = dnpp)) +
  geom_hex(bins = 50, show.legend = FALSE) +
  theme_classic() +
  geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 0, linetype = "dotted") +
  khroma::scale_fill_batlowW(trans = "log", reverse = TRUE) +
  xlim(-1, 2) + 
  ylim(-1, 2) +
  labs(
    x = expression(paste(Delta, "GPP/GPP")),
    y = expression(paste(Delta, "NPP/NPP"))
  )

# Number for text
print("Quartiles of L_NPP:GPP")
print(format(quantile(gdf_sampled$l_npp_gpp, probs =  c(0.25, 0.5, 0.75), na.rm = TRUE), digits = 3))

# Lables for Histogram  
labels <- gdf_sampled %>%
  ungroup()%>%
  mutate(l_larger = l_npp_gpp > 1) %>%
  summarise(
    n_recs = n(),
    perc_larger = round(sum(l_larger, na.rm = TRUE)/n_recs, digits = 3),
    perc_smaller = 1 - perc_larger
    )

# Density plot
gg_density <- gdf_sampled %>% 
  ggplot(aes(x = l_npp_gpp, y = after_stat(density))) +
  geom_density(fill = "grey70") +
  theme_classic() +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 1, linetype = "dotted") +
  xlim(-1, 2.5) + 
  geom_text(
    data = labels,
    aes(x = 0.5, y = Inf, label = perc_smaller),
    vjust = 1.5
    ) +
  geom_text(
    data = labels,
    aes(x = 1.5, y = Inf, label = perc_larger),
    vjust = 1.5
    ) +
  labs(
    x = expression(paste(italic(L)[NPP:GPP]))
  )

# Combine density scatter and density distribution plots
gg3_gpp_npp <- cowplot::plot_grid(
  gg_densityscatter, 
  gg_density, 
  align = 'h',     # 'v' for vertical alignment
  axis = 'tb',
  ncol = 2,
  labels = c("a", "b")
)

gg3_gpp_npp

#ggsave(here("fig/dgpp_dnpp_trendy_s1_pooled.pdf"), plot = gg3_gpp_npp, width = 8, height = 4)
```

#### By model

```{r gpp_npp_bymodl, fig.height=12, out.width="100%"}
set.seed(1982)

gdf_sampled <- gdf %>%
  unnest(data) %>%
  bind_rows(
    .,
    gdf %>%
      mutate(n = purrr::map_int(data, ~nrow(.))) %>%
      mutate(data = purrr::map(data, ~slice_sample(., n = nsample, replace = TRUE))) %>% # smallest set here
      unnest(data) %>%
      mutate(modl = "ALL")
  ) %>%
  dplyr::filter(
    !is.nan(dnpp), !is.nan(dgpp), 
    !is.infinite(dnpp), !is.infinite(dgpp),
    !is.nan(l_npp_gpp), !is.infinite(l_npp_gpp))
gdf_sampled$modl <- factor(gdf_sampled$modl, levels = model_order_with_all)
# density scatter
gdf_sampled %>%
  ggplot(aes(x = dgpp, y = dnpp)) +
  geom_hex(bins = 50, show.legend = FALSE) +
  theme_classic() +
  geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
  geom_hline(yintercept = 0, linetype = "dotted") +
  khroma::scale_fill_batlowW(trans = "log", reverse = TRUE) +
  xlim(-1, 2) + 
  ylim(-1, 2) +
  labs(
    x = expression(paste(Delta, "GPP/GPP")),
    y = expression(paste(Delta, "NPP/NPP"))
  ) +
  facet_wrap(. ~ modl,  ncol = 3,  scales = "free",labeller = as_labeller(display_titles_with_all)) +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(size = 12, hjust = 0)
    )

ggsave(here("fig/dnpp_dgpp_trendy_s1_bymodl.pdf"), width = 9, height = 13)
ggsave(here("fig/dnpp_dgpp_trendy_s1_bymodl.png"), width = 9, height = 13)

# we need to get some stat for the overall excel table
df_quantiles <- gdf_sampled %>%
  group_by(modl) %>%
  summarise(
    l_npp_gpp_50 = quantile(l_npp_gpp, probs = 0.5, na.rm = TRUE),
    l_npp_gpp_iqr = IQR(l_npp_gpp, na.rm = TRUE)
    ) 

summary_table <- left_join(summary_table, df_quantiles, by = 'modl')

# Lables for Histogram  
labels <- gdf_sampled %>%
  mutate(l_larger = l_npp_gpp > 1) %>%
  group_by(modl) %>%
  summarise(
    n_recs = n(),
    perc_larger = round(sum(l_larger, na.rm = TRUE)/n_recs,digits = 2),
    perc_smaller = 1 - perc_larger
    )

# Histograms per model
gdf_sampled %>%
  ggplot() +
  geom_density(aes(x = l_npp_gpp, y = after_stat(density)),fill = "grey70") +
  theme_classic() +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 1, linetype = "dotted") +
  xlim(-1, 2.5) +
  geom_text(data = labels, aes(x = 0.5, y = Inf, label = perc_smaller), vjust =  1.5) +
  geom_text(data = labels, aes(x = 1.5, y = Inf, label = perc_larger), vjust = 1.5)+
  labs(
    x = expression(paste(italic(L)[NPP:GPP]))
  ) +
  facet_wrap(. ~ modl, ncol = 3,  scales = "free",labeller = as_labeller(display_titles_with_all))

ggsave(here("fig/hist_dnpp_dgpp_trendy_bymodl.pdf"), width = 9, height = 10.5)
ggsave(here("fig/hist_dnpp_dgpp_trendy_bymodl.png"), width = 9, height = 10.5)
```


#### Maps

By model.
```{r gpp_npp_bymodl_map, fig.height=12, out.width="100%"}
plot_by_model <- function(use_modl, df){
  
  gdf_long <- df %>%
    unnest(data) %>%
    dplyr::filter(!is.nan(l_npp_gpp), !is.infinite(l_npp_gpp)) |> 
    filter(modl == use_modl)
  
  ggplot() +
    geom_tile(
      data = gdf_long,
      aes(x = lon, y = lat, fill = l_npp_gpp),
      show.legend = TRUE
      ) +
    geom_sf(
      data = coast,
      colour = 'black',
      linewidth = 0.3
    )  +
    coord_sf(
      ylim = c(-60, 85),
      expand = FALSE
    ) +
    khroma::scale_fill_vik(
      name = expression(paste(italic(L)[NPP:GPP])),
      limits = c(0, 2), 
      oob = scales::squish,
      midpoint = 1,
      reverse = TRUE
      ) +  # suppress individual legends
    labs(
      title = display_titles[use_modl]
    ) +
    theme_void() +
    theme(
      legend.position = "none",
      plot.title = element_text(hjust = 0.5, size = 10) 
      )
}

plotlist <- purrr::map(
  as.list(unique(gdf$modl)),
  ~ plot_by_model(., gdf)
)

legend_plot <- plotlist[[1]] +
    theme(legend.position = "right")

legend <- cowplot::get_legend(legend_plot)

map_grid <- cowplot::plot_grid(
  plotlist = plotlist,
  ncol = 3
)

final_plot <- cowplot::plot_grid(
  map_grid,
  legend,
  ncol = 2,
  rel_widths = c(1, 0.08)  # adjust the legend width as needed
)

final_plot

# ggsave(
#   filename = here("fig/map_dgpp_dnpp_bymodel.pdf"),
#   height = 10,
#   width = 10
#   )
#  save the old version for revision purpose

ggsave(
  filename = here("fig/map_dgpp_dnpp_bymodelv2.pdf"),
  height = 10,
  width = 10
  )

ggsave(
  filename = here("fig/map_dgpp_dnpp_bymodelv2.png"),
  height = 10,
  width = 10
  )
```

All models pooled.
```{r}
gdf_long_1x1deg <- gdf %>%
  select(-data) %>% 
  unnest(data_1x1deg) %>%
  group_by(lon, lat) %>% 
  summarise(
    l_npp_gpp_median = median(l_npp_gpp, na.rm = TRUE),
    l_npp_gpp_sd = sd(l_npp_gpp, na.rm = TRUE),
    nmodels = sum(!is.na(l_npp_gpp)),.groups = "drop"
  )%>%
  mutate(across(
    .cols = -c(lon, lat, nmodels), 
    .fns = ~ ifelse(nmodels < 8, NA, .x)
  ))
  
gg_map_agg <- ggplot() +
  geom_tile(
    data = gdf_long_1x1deg,
    aes(x = lon, y = lat, fill = l_npp_gpp_median),
    show.legend = TRUE
    ) +
  geom_sf(
    data = coast,
    colour = 'black',
    linewidth = 0.3
  )  +
  coord_sf(
    ylim = c(-60, 85),
    expand = FALSE
  ) +
  khroma::scale_fill_vik(
    name = expression(atop('Median of',italic(L)[NPP:GPP])),
    limits = c(0.2, 1.8),  
    oob = scales::squish,
    midpoint = 1,
    reverse = TRUE
    ) +
  theme_void()

gg_map_agg

gg_map_aggSD <- ggplot() +
  geom_tile(
    data = gdf_long_1x1deg,
    aes(x = lon, y = lat, fill = l_npp_gpp_sd),
    show.legend = TRUE
    ) +
  geom_sf(
    data = coast,
    colour = 'black',
    linewidth = 0.3
  )  +
  coord_sf(
    ylim = c(-60, 85),
    expand = FALSE
  ) +
  khroma::scale_fill_grayC(
    name = expression(atop('SD of',italic(L)[NPP:GPP])),
    limits = c(0, 2),
    oob = scales::squish
    ) +
  theme_void()

gg_map_aggSD



```

#### Publication Fig. 1

```{r}
cowplot::plot_grid(
  gg3_gpp_npp,
  gg_map_agg,
  gg_map_aggSD,
  ncol = 1,
  labels = c("", "c", "d"),
  rel_heights = c(1, 0.9,0.9)
)

ggsave(filename = here("fig/fig1_combined_npp_gpp_sd.png"), height = 8, width = 6)
ggsave(filename = here("fig/fig1_combined_npp_gpp_sd.pdf"), height = 8, width = 6)
``` 

### Cveg* vs NPP 

#### All models pooled

```{r}
set.seed(1982)

gdf_sampled <- gdf %>%
  mutate(data = purrr::map(data, ~slice_sample(., n = nsample, replace = TRUE))) %>%
  unnest(data) %>%
  dplyr::filter(!is.nan(dnpp), !is.nan(dcveg_star), !is.infinite(dnpp), !is.infinite(dcveg_star),l_cveg_star_npp>-1000000,l_cveg_star_npp<1000000) # the infinite filter does not work, no idea why

# density scatter-plot for pooled data, balanced samples, original grid
gg_densityscatter <- gdf_sampled %>%
  ggplot(aes(x = dnpp, y = dcveg_star)) +
  geom_hex(bins = 50, show.legend = FALSE) +
  theme_classic() +
  geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 0, linetype = "dotted") +
  khroma::scale_fill_batlowW(trans = "log", reverse = TRUE) +
  xlim(-1, 2) + 
  ylim(-1, 2) +
  labs(
    x = expression(paste(Delta, "NPP/NPP")),
    y = expression(paste(Delta, italic(C)[veg^"*"], "/", italic(C)[veg]))
  )

# Number for text
print("Quartiles of L_Cveg:NPP")
print(format(quantile(gdf_sampled$l_cveg_star_npp, probs =  c(0.25, 0.5, 0.75), na.rm = TRUE), digits = 3))

# Lables for Histogram  
labels <- gdf_sampled %>%
  ungroup()%>%
  mutate(l_larger = l_cveg_star_npp > 1) %>%
  summarise(
    n_recs = n(),
    perc_larger = round(sum(l_larger, na.rm = TRUE)/n_recs, digits = 3),
    perc_smaller = 1 - perc_larger
    )

# Density plot
gg_density <- gdf_sampled %>% 
  ggplot(aes(x = l_cveg_star_npp, y = after_stat(density))) +
  geom_density(fill = "grey70") +
  theme_classic() +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 1, linetype = "dotted") +
  xlim(-1, 2.5) + 
  geom_text(
    data = labels,
    aes(x = 0.5, y = Inf, label = perc_smaller),
    vjust = 1.5) +
  geom_text(
    data = labels,
    aes(x = 1.5, y = Inf, label = perc_larger),
    vjust = 1.5) +
  labs(
    x = expression(paste(italic(L)[Cveg^"*":NPP]))
  )

# Combine density scatter and density distribution plots
gg3_cveg_star_npp <- cowplot::plot_grid(
  gg_densityscatter, 
  gg_density, 
  align = 'h',     # 'v' for vertical alignment
  axis = 'tb',
  ncol = 2,
  labels = c("a", "b")
)

#ggsave(here("fig/cveg_star_npp_trendy_pooled.pdf"), plot = gg3_cveg_star_npp, width = 8, height = 4)

gg3_cveg_star_npp
```

#### By model

```{r npp_cvegstar_bymodl, fig.height=12, out.width="100%"}
set.seed(1982)

gdf_sampled <- gdf %>%
  unnest(data) %>%
  bind_rows(
    .,
    gdf %>%
      mutate(n = purrr::map_int(data, ~nrow(.))) %>%
      mutate(data = purrr::map(data, ~slice_sample(., n = nsample, replace = TRUE))) %>% # smallest set here
      unnest(data) %>%
      mutate(modl = "ALL")
  ) %>%
  dplyr::filter(!is.nan(dnpp), !is.nan(dcveg_star), !is.infinite(dnpp), !is.infinite(dcveg_star),l_cveg_star_npp>-1000000,l_cveg_star_npp<1000000) # the infinite filter does not work, no idea why
gdf_sampled$modl <- factor(gdf_sampled$modl, levels = model_order_with_all)

# density scatter
gdf_sampled %>%
  ggplot(aes(x = dnpp, y = dcveg_star)) +
  geom_hex(bins = 50, show.legend = FALSE) +
  theme_classic() +
  geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
  geom_hline(yintercept = 0, linetype = "dotted") +
  # geom_vline(xintercept = 0, linetype = "dotted") +
  khroma::scale_fill_batlowW(trans = "log", reverse = TRUE) +
  xlim(-1, 2) + 
  ylim(-1, 2) +
  labs(
    x = expression(paste(Delta, "NPP/NPP")),
    y = expression(paste(Delta, italic(C)[veg^"*"], "/", italic(C)[veg]))
  ) +
  facet_wrap(. ~ modl,  ncol = 3,  scales = "free",labeller = as_labeller(display_titles_with_all)) +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(size = 12, hjust = 0)
    )

ggsave(here("fig/dnpp_dvegc_trendy_s1_bymodl.pdf"), width = 9, height = 13)
ggsave(here("fig/dnpp_dvegc_trendy_s1_bymodl.png"), width = 9, height = 13)

# we need to get some stat for the overall excel table
df_quantiles <- gdf_sampled %>%
  group_by(modl) %>%
  summarise(
    l_cveg_star_npp_50 = quantile(l_cveg_star_npp, probs = 0.5, na.rm = TRUE),
    l_cveg_star_npp_iqr = IQR(l_cveg_star_npp, na.rm = TRUE)
    ) 

summary_table <- left_join(summary_table, df_quantiles, by = 'modl')

# Lables for Histogram  
labels <- gdf_sampled %>%
  mutate(l_larger = l_cveg_star_npp > 1) %>%
  group_by(modl) %>%
  summarise(
    n_recs = n(),
    perc_larger = round(sum(l_larger, na.rm = TRUE)/n_recs,digits = 2),
    perc_smaller = 1 - perc_larger
    )

# Histograms per model
gdf_sampled %>%
  ggplot() +
  geom_density(aes(x = l_cveg_star_npp, y = after_stat(density)), fill = "grey70") +
  theme_classic() +
  geom_hline(yintercept = 0, linetype = "dotted") +
  # geom_vline(xintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 1, linetype = "dotted") +
  xlim(-1, 2.5) +
  geom_text(data = labels, aes(x = 0.5, y = Inf, label = perc_smaller), vjust =  1.5) +
  geom_text(data = labels, aes(x = 1.5, y = Inf, label = perc_larger), vjust = 1.5)+
  labs(
    x = expression(paste(italic(L)[Cveg^"*":NPP]))
  ) +
  facet_wrap(. ~ modl, ncol = 3,  scales = "free",labeller = as_labeller(display_titles_with_all)) 

ggsave(here("fig/hist_dcveg_dnpp_trendy_bymodl.pdf"), width = 9, height = 10.5)
ggsave(here("fig/hist_dcveg_dnpp_trendy_bymodl.png"), width = 9, height = 10.5)
```

#### Maps

By model.
```{r cveg_npp_bymodl_map, fig.height=12, out.width="100%"}
plot_by_model <- function(use_modl, df){
  
  gdf_long <- df %>%
    unnest(data) %>%
    dplyr::filter(!is.nan(l_cveg_star_npp), !is.infinite(l_cveg_star_npp)) |> 
    filter(modl == use_modl)
  
  ggplot() +
    geom_tile(
      data = gdf_long,
      aes(x = lon, y = lat, fill = l_cveg_star_npp),
      show.legend = TRUE
      ) +
    geom_sf(
      data = coast,
      colour = 'black',
      linewidth = 0.3
    )  +
    coord_sf(
      ylim = c(-60, 85),
      expand = FALSE
    ) +
    khroma::scale_fill_vik(
      name = expression(paste(italic(L)[Cveg^"*":NPP])),
      limits = c(0, 2), 
      oob = scales::squish,
      midpoint = 1,
      reverse = TRUE
      ) +
    labs(
      title = display_titles[use_modl]
    ) +
    theme_void() +
    theme(
      legend.position = "none",
      plot.title = element_text(hjust = 0.5, size = 10) 
      )
}

plotlist <- purrr::map(
  model_order,
  ~ plot_by_model(., gdf)
)

legend_plot <- plotlist[[1]] +
    theme(legend.position = "right")

legend <- cowplot::get_legend(legend_plot)

map_grid <- cowplot::plot_grid(
  plotlist = plotlist,
  ncol = 3
)

final_plot <- cowplot::plot_grid(
  map_grid,
  legend,
  ncol = 2,
  rel_widths = c(1, 0.08)  # adjust the legend width as needed
)

final_plot

ggsave(
  filename = here("fig/map_dcvegstar_dnpp_bymodel.pdf"),
  height = 10, 
  width = 10
  )

ggsave(
  filename = here("fig/map_dcvegstar_dnpp_bymodel.png"),
  height = 10, 
  width = 10
  )
```

All models pooled.
```{r}
gdf_long_1x1deg <- gdf %>%
  select(-data) %>% 
  unnest(data_1x1deg) %>%
  group_by(lon, lat) %>% 
  summarise(
    l_cveg_star_npp_median = median(l_cveg_star_npp, na.rm = TRUE),
    l_cveg_star_npp_sd = sd(l_cveg_star_npp, na.rm = TRUE),
    nmodels = sum(!is.na(l_cveg_star_npp)),.groups = "drop"
  )%>%
  mutate(across(
    .cols = -c(lon, lat, nmodels), 
    .fns = ~ ifelse(nmodels < 8, NA, .x)
  ))
  
gg_map_agg <- ggplot() +
  geom_tile(
    data = gdf_long_1x1deg,
    aes(x = lon, y = lat, fill = l_cveg_star_npp_median),
    show.legend = TRUE
    ) +
  geom_sf(
    data = coast,
    colour = 'black',
    linewidth = 0.3
  )  +
  coord_sf(
    ylim = c(-60, 85),
    expand = FALSE
  ) +
  khroma::scale_fill_vik(
    name = expression(atop('Median of',italic(L)[Cveg^"*":NPP])),
    limits = c(0.2, 1.8),  
    midpoint = 1,
    oob = scales::squish,
    reverse = TRUE
    ) +
  theme_void()

gg_map_agg


gg_map_aggSD <- ggplot() +
  geom_tile(
    data = gdf_long_1x1deg,
    aes(x = lon, y = lat, fill = l_cveg_star_npp_sd),
    show.legend = TRUE
    ) +
  geom_sf(
    data = coast,
    colour = 'black',
    linewidth = 0.3
  )  +
  coord_sf(
    ylim = c(-60, 85),
    expand = FALSE
  ) +
  khroma::scale_fill_grayC(
    name = expression(atop("SD of", italic(L)[Cveg^"*":NPP])),
    limits = c(0, 2),
    oob = scales::squish
  ) +
  theme_void()

gg_map_aggSD
```

#### Publication Fig. 3

```{r}
cowplot::plot_grid(
  gg3_cveg_star_npp,
  gg_map_agg,
  gg_map_aggSD,
  ncol = 1,
  labels = c("", "c", "d"),
  rel_heights = c(1, 0.9,0.9)
)

ggsave(filename = here("fig/fig3_combined_cvegstar_npp_sd.png"), height = 8, width = 6)
ggsave(filename = here("fig/fig3_combined_cvegstar_npp_sd.pdf"), height = 8, width = 6)
``` 



### Cveg vs NPP 

#### All models pooled (similar to the above but no star)

```{r}
set.seed(1982)

gdf_sampled <- gdf %>%
  mutate(data = purrr::map(data, ~slice_sample(., n = nsample, replace = TRUE))) %>%
  unnest(data) %>%
  mutate(l_cveg_npp = dcveg / dnpp)%>%
  dplyr::filter(!is.nan(dnpp), !is.nan(dcveg), !is.infinite(dnpp), !is.infinite(dcveg),l_cveg_npp>-1000000,l_cveg_npp<1000000) # the infinite filter does not work, no idea why

# density scatter-plot for pooled data, balanced samples, original grid
gg_densityscatter <- gdf_sampled %>%
  ggplot(aes(x = dnpp, y = dcveg)) +
  geom_hex(bins = 50, show.legend = FALSE) +
  theme_classic() +
  geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 0, linetype = "dotted") +
  khroma::scale_fill_batlowW(trans = "log", reverse = TRUE) +
  xlim(-1, 2) + 
  ylim(-1, 2) +
  labs(
    x = expression(paste(Delta, "NPP/NPP")),
    y = expression(paste(Delta, italic(C)[veg], "/", italic(C)[veg]))
  )

# Number for text
print("Quartiles of L_Cveg:NPP")
print(format(quantile(gdf_sampled$l_cveg_npp, probs =  c(0.25, 0.5, 0.75), na.rm = TRUE), digits = 3))

# Lables for Histogram  
labels <- gdf_sampled %>%
  ungroup()%>%
  mutate(l_larger = l_cveg_npp > 1) %>%
  summarise(
    n_recs = n(),
    perc_larger = round(sum(l_larger, na.rm = TRUE)/n_recs, digits = 2),
    perc_smaller = 1 - perc_larger
    )

# Density plot
gg_density <- gdf_sampled %>% 
  ggplot(aes(x = l_cveg_npp, y = after_stat(density))) +
  geom_density(fill = "grey70") +
  theme_classic() +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 1, linetype = "dotted") +
  xlim(-1, 2.5) + 
  geom_text(
    data = labels,
    aes(x = 0.5, y = Inf, label = perc_smaller),
    vjust = 1.5) +
  geom_text(
    data = labels,
    aes(x = 1.5, y = Inf, label = perc_larger),
    vjust = 1.5) +
  labs(
    x = expression(paste(italic(L)[Cveg:NPP]))
  )

# Combine density scatter and density distribution plots
gg3_cveg_npp <- cowplot::plot_grid(
  gg_densityscatter, 
  gg_density, 
  align = 'h',     # 'v' for vertical alignment
  axis = 'tb',
  ncol = 2,
  labels = c("a", "b")
)

#ggsave(here("fig/cveg_npp_trendy_pooled.pdf"), plot = gg3_cveg_star_npp, width = 8, height = 4)

gg3_cveg_npp
```

#### By model

```{r npp_cveg_bymodl, fig.height=12, out.width="100%"}
set.seed(1982)

gdf_sampled <- gdf %>%
  unnest(data) %>%
  bind_rows(
    .,
    gdf %>%
      mutate(n = purrr::map_int(data, ~nrow(.))) %>%
      mutate(data = purrr::map(data, ~slice_sample(., n = nsample, replace = TRUE))) %>% # smallest set here
      unnest(data) %>%
      mutate(modl = "ALL")
  ) %>%
  mutate(l_cveg_npp = dcveg / dnpp)%>%
  dplyr::filter(!is.nan(dnpp), !is.nan(dcveg), !is.infinite(dnpp), !is.infinite(dcveg), l_cveg_npp>-1000000,l_cveg_npp<1000000) # the infinite filter does not work, no idea why
gdf_sampled$modl <- factor(gdf_sampled$modl, levels = model_order_with_all)

# density scatter
gdf_sampled %>%
  ggplot(aes(x = dnpp, y = dcveg)) +
  geom_hex(bins = 50, show.legend = FALSE) +
  theme_classic() +
  geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
  geom_hline(yintercept = 0, linetype = "dotted") +
  # geom_vline(xintercept = 0, linetype = "dotted") +
  khroma::scale_fill_batlowW(trans = "log", reverse = TRUE) +
  xlim(-1, 2) + 
  ylim(-1, 2) +
  labs(
    x = expression(paste(Delta, "NPP/NPP")),
    y = expression(paste(Delta, italic(C)[veg], "/", italic(C)[veg]))
  ) +
  facet_wrap(. ~ modl,  ncol = 3,  scales = "free",labeller = as_labeller(display_titles_with_all)) +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(size = 12, hjust = 0)
    )

ggsave(here("fig/dnpp_dvegc_trendy_s1_bymodl_nostar.pdf"), width = 9, height = 13)
ggsave(here("fig/dnpp_dvegc_trendy_s1_bymodl_nostar.png"), width = 9, height = 13)

# we need to get some stat for the overall excel table
df_quantiles <- gdf_sampled %>%
  group_by(modl) %>%
  summarise(
    l_cveg_npp_50 = quantile(l_cveg_npp, probs = 0.5, na.rm = TRUE),
    l_cveg_npp_iqr = IQR(l_cveg_npp, na.rm = TRUE)
    ) 

summary_table <- left_join(summary_table, df_quantiles, by = 'modl')

# Lables for Histogram  
labels <- gdf_sampled %>%
  mutate(l_larger = l_cveg_npp > 1) %>%
  group_by(modl) %>%
  summarise(
    n_recs = n(),
    perc_larger = round(sum(l_larger, na.rm = TRUE)/n_recs,digits = 2),
    perc_smaller = 1 - perc_larger
    )

# Histograms per model
gdf_sampled %>%
  ggplot() +
  geom_density(aes(x = l_cveg_npp, y = after_stat(density)), fill = "grey70") +
  theme_classic() +
  geom_hline(yintercept = 0, linetype = "dotted") +
  # geom_vline(xintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 1, linetype = "dotted") +
  xlim(-1, 2.5) +
  geom_text(data = labels, aes(x = 0.5, y = Inf, label = perc_smaller), vjust =  1.5) +
  geom_text(data = labels, aes(x = 1.5, y = Inf, label = perc_larger), vjust = 1.5)+
  labs(
    x = expression(paste(italic(L)[Cveg:NPP]))
  ) +
  facet_wrap(. ~ modl, ncol = 3,  scales = "free",labeller = as_labeller(display_titles_with_all)) 

ggsave(here("fig/hist_dcveg_dnpp_trendy_bymodl_nostar.png"), width = 9, height = 10.5)
```

#### Maps

By model.
```{r cveg_npp_bymodl_map, fig.height=12, out.width="100%"}
plot_by_model <- function(use_modl, df){
  
  gdf_long <- df %>%
    unnest(data) %>%
    mutate(l_cveg_npp = dcveg / dnpp)%>%
    dplyr::filter(!is.nan(l_cveg_npp), !is.infinite(l_cveg_npp)) |> 
    filter(modl == use_modl)
  
  ggplot() +
    geom_tile(
      data = gdf_long,
      aes(x = lon, y = lat, fill = l_cveg_npp),
      show.legend = TRUE
      ) +
    geom_sf(
      data = coast,
      colour = 'black',
      linewidth = 0.3
    )  +
    coord_sf(
      ylim = c(-60, 85),
      expand = FALSE
    ) +
    khroma::scale_fill_vik(
      name = expression(paste(italic(L)[Cveg:NPP])),
      limits = c(0, 2), 
      oob = scales::squish,
      midpoint = 1,
      reverse = TRUE
      ) +
    labs(
      title = display_titles[use_modl]
    ) +
    theme_void() +
    theme(
      legend.position = "none",
      plot.title = element_text(hjust = 0.5, size = 10) 
      )
}

plotlist <- purrr::map(
  model_order,
  ~ plot_by_model(., gdf)
)

legend_plot <- plotlist[[1]] +
    theme(legend.position = "right")

legend <- cowplot::get_legend(legend_plot)

map_grid <- cowplot::plot_grid(
  plotlist = plotlist,
  ncol = 3
)

final_plot <- cowplot::plot_grid(
  map_grid,
  legend,
  ncol = 2,
  rel_widths = c(1, 0.08)  # adjust the legend width as needed
)

final_plot


ggsave(
  filename = here("fig/map_dcveg_dnpp_bymodel_nostar.png"),
  height = 10, 
  width = 10
  )
```

All models pooled.
```{r}
gdf_long_1x1deg <- gdf %>%
  select(-data) %>% 
  unnest(data_1x1deg) %>%
  mutate(l_cveg_npp = dcveg / dnpp)%>%
  group_by(lon, lat) %>% 
  summarise(
    l_cveg_npp_median = median(l_cveg_npp, na.rm = TRUE),
    l_cveg_npp_sd = sd(l_cveg_npp, na.rm = TRUE),
    nmodels=n()
  )
  
gg_map_agg <- ggplot() +
  geom_tile(
    data = gdf_long_1x1deg,
    aes(x = lon, y = lat, fill = l_cveg_npp_median),
    show.legend = TRUE
    ) +
  geom_sf(
    data = coast,
    colour = 'black',
    linewidth = 0.3
  )  +
  coord_sf(
    ylim = c(-60, 85),
    expand = FALSE
  ) +
  khroma::scale_fill_vik(
    name = expression(atop('Median of',italic(L)[Cveg:NPP])),
    limits = c(0.2, 1.8),  
    midpoint = 1,
    oob = scales::squish,
    reverse = TRUE
    ) +
  theme_void()

gg_map_agg


gg_map_aggSD <- ggplot() +
  geom_tile(
    data = gdf_long_1x1deg,
    aes(x = lon, y = lat, fill = l_cveg_npp_sd),
    show.legend = TRUE
    ) +
  geom_sf(
    data = coast,
    colour = 'black',
    linewidth = 0.3
  )  +
  coord_sf(
    ylim = c(-60, 85),
    expand = FALSE
  ) +
  khroma::scale_fill_vik(
    name = expression(atop('SD of',italic(L)[Cveg:NPP])),
    limits = c(0, 2),  
    oob = scales::squish,
    midpoint = 1,
    reverse = TRUE
    ) +
  theme_void()

gg_map_aggSD
```

#### SI Fig. XXX

```{r}
cowplot::plot_grid(
  gg3_cveg_npp,
  gg_map_agg,
  gg_map_aggSD,
  ncol = 1,
  labels = c("", "c", "d"),
  rel_heights = c(1, 0.9,0.9)
)

ggsave(filename = here("fig/combined_cveg_npp_sd_nostar.png"), height = 8, width = 6)
ggsave(filename = here("fig/combined_cveg_npp_sd_nostar.pdf"), height = 8, width = 6)
``` 




### Croot vs Cveg

Taking the ratio of Croot vs. Cveg here, both for their non-steady conditions (without applying Eq. 12 from paper). This is because root C definition is inconsistent across models and may include (or not) coarse roots with slow turnover - similar to that of wood C (reflected in total vegetation C).  

#### All models pooled

```{r}
set.seed(1982)

gdf_sampled <- gdf %>%
  mutate(data = purrr::map(data, ~slice_sample(., n = nsample))) %>%
  unnest(data) %>%
  dplyr::filter(
   !is.nan(dcroot), !is.nan(dcveg), 
    !is.infinite(dcroot), !is.infinite(dcveg),
    !is.nan(l_croot_cveg), !is.infinite(l_croot_cveg))

# density scatter-plot for pooled data, balanced samples, original grid
gg_densityscatter <- gdf_sampled %>%
  ggplot(aes(x = dcveg, y = dcroot)) +
  geom_hex(bins = 50, show.legend = FALSE) +
  theme_classic() +
  geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 0, linetype = "dotted") +
  khroma::scale_fill_batlowW(trans = "log", reverse = TRUE) +
  xlim(-1, 2) + 
  ylim(-1, 2) +
  labs(
    x = expression(paste(Delta, italic(C)[veg], "/", italic(C)[veg])),
    y = expression(paste(Delta, italic(C)[root], "/", italic(C)[root])),
  )

# Number for text
print("Quartiles of L_Croot:Cveg")
print(format(quantile(gdf_sampled$l_croot_cveg, probs =  c(0.25, 0.5, 0.75), na.rm = TRUE), digits = 3))

# Lables for Histogram  
labels <- gdf_sampled %>%
  ungroup()%>%
  mutate(l_larger = l_croot_cveg > 1) %>%
  summarise(
    n_recs = n(),
    perc_larger = round(sum(l_larger, na.rm = TRUE)/n_recs,digits = 3),
    perc_smaller = 1 - perc_larger
    )

# Density plot
gg_density <- gdf_sampled %>% 
  ggplot(aes(x = l_croot_cveg, y = after_stat(density))) +
  geom_density(fill = "grey70") +
  theme_classic() +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 1, linetype = "dotted") +
  xlim(-1, 2.5) + 
  geom_text(
    data = labels,
    aes(x = 0.5, y = Inf, label = perc_smaller),
    vjust= 1.5) +
  geom_text(
    data = labels,
    aes(x = 1.5, y = Inf, label = perc_larger),
    vjust= 1.5) +
  labs(
    x = expression(paste(italic(L)[Croot:Cveg]))
  )

# Combine density scatter and density distribution plots
gg3_l_croot_cveg <- cowplot::plot_grid(
  gg_densityscatter, 
  gg_density, 
  align = 'h',     # 'v' for vertical alignment
  axis = 'tb',
  ncol = 2,
  labels = c("a", "b")
)

# ggsave(here("fig/croot_cveg_pooled.pdf"), plot = gg3_l_croot_cveg, width = 8, height = 4)
```

#### sensitivity test using JULES only

```{r cveg_star_croot_bymodl, fig.height=9, out.width="100%"}

set.seed(1982)

model_no <- which(gdf$modl=="JULES") 
model_df <- gdf$data[[model_no]]
robustbase::colMedians(as.matrix(model_df), na.rm = TRUE)
sum(!model_df$landCoverFrac==0)

gdf_sampled <- gdf %>%
  unnest(data) %>%
  filter(modl == "JULES")%>%
  mutate(l_croot_star_cveg_star=dcroot_star/dcveg_star)%>%
  dplyr::filter(
    !is.nan(dcroot_star), !is.nan(dcveg_star), 
    !is.infinite(dcroot_star), !is.infinite(dcveg_star),
        !is.nan(dcroot), !is.nan(dcveg), 
    !is.infinite(dcroot), !is.infinite(dcveg),
    !is.nan(l_croot_cveg), !is.infinite(l_croot_cveg),
    !is.nan(l_croot_star_cveg_star), !is.infinite(l_croot_star_cveg_star),
    !is.nan(l_croot_cveg_star), !is.infinite(l_croot_cveg_star))

# density scatter

double_stara<-gdf_sampled %>%
  ggplot(aes(x = dcveg_star, y = dcroot_star)) +
  geom_hex(bins = 50, show.legend = FALSE) +
  theme_classic() +
  geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 0, linetype = "dotted") +
  khroma::scale_fill_batlowW(trans = "log", reverse = TRUE) +
  xlim(-1, 2) + 
  ylim(-1, 2) +
  labs(
    x = expression(paste(Delta, italic(C)[veg^"*"], "/", italic(C)[veg])),
    y = expression(paste(Delta, italic(C)[root^"*"], "/", italic(C)[root]))
  ) +
  facet_wrap(. ~ modl,  ncol = 3,  scales = "free") +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(size = 12, hjust = 0)
    )

one_stara<-gdf_sampled %>%
  ggplot(aes(x = dcveg_star, y = dcroot)) +
  geom_hex(bins = 50, show.legend = FALSE) +
  theme_classic() +
  geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 0, linetype = "dotted") +
  khroma::scale_fill_batlowW(trans = "log", reverse = TRUE) +
  xlim(-1, 2) + 
  ylim(-1, 2) +
  labs(
    x = expression(paste(Delta, italic(C)[veg^"*"], "/", italic(C)[veg])),
    y = expression(paste(Delta, italic(C)[root], "/", italic(C)[root]))
  ) +
  facet_wrap(. ~ modl,  ncol = 3,  scales = "free") +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(size = 12, hjust = 0)
    )


no_stara<-gdf_sampled %>%
  ggplot(aes(x = dcveg, y = dcroot)) +
  geom_hex(bins = 50, show.legend = FALSE) +
  theme_classic() +
  geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 0, linetype = "dotted") +
  khroma::scale_fill_batlowW(trans = "log", reverse = TRUE) +
  xlim(-1, 2) + 
  ylim(-1, 2) +
  labs(
    x = expression(paste(Delta, italic(C)[veg], "/", italic(C)[veg])),
    y = expression(paste(Delta, italic(C)[root], "/", italic(C)[root]))
  ) +
  facet_wrap(. ~ modl,  ncol = 3,  scales = "free") +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(size = 12, hjust = 0)
    )


# Lables for Histogram  
labels <- gdf_sampled %>%
  mutate(l_larger = l_croot_star_cveg_star > 1) %>%
  group_by(modl) %>%
  summarise(
    n_recs = n(),
    perc_larger = round(sum(l_larger, na.rm = TRUE)/n_recs, digits = 2),
    perc_smaller = 1 - perc_larger
    )

# Histograms per model
double_star<-gdf_sampled %>%
  ggplot() +
  geom_density(aes(x = l_croot_star_cveg_star, y = after_stat(density)), fill = "grey70") +
  theme_classic() +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 1, linetype = "dotted") +
  xlim(-1, 2.5) +
  geom_text(data = labels, aes(x = 0.5, y = Inf, label = perc_smaller), vjust =  1.5) +
  geom_text(data = labels, aes(x = 1.5, y = Inf, label = perc_larger), vjust = 1.5)+
  labs(
    x = expression(paste(italic(L)[Croot^"*":Cveg^"*"]))
  ) +
  facet_wrap(. ~ modl, ncol = 3,  scales = "free") 

# Lables for Histogram  
labels <- gdf_sampled %>%
  mutate(l_larger = l_croot_cveg_star > 1) %>%
  group_by(modl) %>%
  summarise(
    n_recs = n(),
    perc_larger = round(sum(l_larger, na.rm = TRUE)/n_recs, digits = 2),
    perc_smaller = 1 - perc_larger
    )

# Histograms per model
one_star<-gdf_sampled %>%
  ggplot() +
  geom_density(aes(x = l_croot_cveg_star, y = after_stat(density)), fill = "grey70") +
  theme_classic() +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 1, linetype = "dotted") +
  xlim(-1, 2.5) +
  geom_text(data = labels, aes(x = 0.5, y = Inf, label = perc_smaller), vjust =  1.5) +
  geom_text(data = labels, aes(x = 1.5, y = Inf, label = perc_larger), vjust = 1.5)+
  labs(
    x = expression(paste(italic(L)[Croot:Cveg^"*"]))
  ) +
  facet_wrap(. ~ modl, ncol = 3,  scales = "free") 


# Lables for Histogram  
labels <- gdf_sampled %>%
  mutate(l_larger = l_croot_cveg > 1) %>%
  group_by(modl) %>%
  summarise(
    n_recs = n(),
    perc_larger = round(sum(l_larger, na.rm = TRUE)/n_recs, digits = 2),
    perc_smaller = 1 - perc_larger
    )

# Histograms per model
no_star<-gdf_sampled %>%
  ggplot() +
  geom_density(aes(x = l_croot_cveg, y = after_stat(density)), fill = "grey70") +
  theme_classic() +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 1, linetype = "dotted") +
  xlim(-1, 2.5) +
  geom_text(data = labels, aes(x = 0.5, y = Inf, label = perc_smaller), vjust =  1.5) +
  geom_text(data = labels, aes(x = 1.5, y = Inf, label = perc_larger), vjust = 1.5)+
  labs(
    x = expression(paste(italic(L)[Croot:Cveg]))
  ) +
  facet_wrap(. ~ modl, ncol = 3,  scales = "free") 

 library(cowplot)
 plot_grid(double_star,double_stara, one_star,one_stara, no_star,no_stara, ncol = 2,labels = "AUTO")
ggsave(filename = here('fig/sensitivity_test_star_croot.png'),width=7,height=9)

```

#### By model

```{r cveg_croot_bymodl, fig.height=9, out.width="100%"}
set.seed(1982)

gdf_sampled <- gdf %>%
  unnest(data) %>%
  bind_rows(
    .,
    gdf %>%
      mutate(n = purrr::map_int(data, ~nrow(.))) %>%
      mutate(data = purrr::map(data, ~slice_sample(., n = nsample))) %>% # smallest set here
      unnest(data) %>%
      mutate(modl = "ALL")
  ) %>%
  dplyr::filter(
    !is.nan(dcroot), !is.nan(dcveg), 
    !is.infinite(dcroot), !is.infinite(dcveg),
    !is.nan(l_croot_cveg), !is.infinite(l_croot_cveg))
gdf_sampled$modl <- factor(gdf_sampled$modl, levels = model_order_with_all)

# density scatter
gdf_sampled %>%
  ggplot(aes(x = dcveg, y = dcroot)) +
  geom_hex(bins = 50, show.legend = FALSE) +
  theme_classic() +
  geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 0, linetype = "dotted") +
  khroma::scale_fill_batlowW(trans = "log", reverse = TRUE) +
  xlim(-1, 2) + 
  ylim(-1, 2) +
  labs(
    x = expression(paste(Delta, italic(C)[veg], "/", italic(C)[veg])),
    y = expression(paste(Delta, italic(C)[root], "/", italic(C)[root]))
  ) +
  facet_wrap(. ~ modl,  ncol = 3,  scales = "free",labeller = as_labeller(display_titles_with_all)) +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(size = 12, hjust = 0)
    )

ggsave(here("fig/dcroot_dcveg_bymodl.pdf"), width = 9, height = 13)
ggsave(here("fig/dcroot_dcveg_bymodl.png"), width = 9, height = 13)

# we need to get some stat for the overall excel table
df_quantiles <- gdf_sampled %>%
  group_by(modl) %>%
  summarise(
    l_croot_cveg_50 = quantile(l_croot_cveg, probs = 0.5, na.rm = TRUE),
    l_croot_cveg_iqr = IQR(l_croot_cveg, na.rm = TRUE)
    ) 

summary_table <- left_join(summary_table, df_quantiles, by = 'modl')

# Lables for Histogram  
labels <- gdf_sampled %>%
  mutate(l_larger = l_croot_cveg > 1) %>%
  group_by(modl) %>%
  summarise(
    n_recs = n(),
    perc_larger = round(sum(l_larger, na.rm = TRUE)/n_recs, digits = 2),
    perc_smaller = 1 - perc_larger
    )

# Histograms per model
gdf_sampled %>%
  ggplot() +
  geom_density(aes(x = l_croot_cveg, y = after_stat(density)), fill = "grey70") +
  theme_classic() +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 1, linetype = "dotted") +
  xlim(-1, 2.5) +
  geom_text(data = labels, aes(x = 0.5, y = Inf, label = perc_smaller), vjust =  1.5) +
  geom_text(data = labels, aes(x = 1.5, y = Inf, label = perc_larger), vjust = 1.5)+
  labs(
    x = expression(paste(italic(L)[Croot:Cveg]))
  ) +
  facet_wrap(. ~ modl, ncol = 3,  scales = "free",labeller = as_labeller(display_titles_with_all)) 

ggsave(here("fig/hist_dcroot_dcveg_bymodl.pdf"), width = 9, height = 10.5)
ggsave(here("fig/hist_dcroot_dcveg_bymodl.png"), width = 9, height = 10.5)
```

#### Maps

By model.
```{r cveg_croot_bymodl_map, fig.height=9, out.width="100%"}
plot_by_model <- function(use_modl, df){
  
  gdf_long <- df %>%
    unnest(data) %>%
    dplyr::filter(!is.nan(l_croot_cveg), !is.infinite(l_croot_cveg)) |> 
    filter(modl == use_modl)
  
  ggplot() +
    geom_tile(
      data = gdf_long,
      aes(x = lon, y = lat, fill = l_croot_cveg),
      show.legend = TRUE
      ) +
    geom_sf(
      data = coast,
      colour = 'black',
      linewidth = 0.3
    )  +
    coord_sf(
      ylim = c(-60, 85),
      expand = FALSE
    ) +
    khroma::scale_fill_vik(
      name = expression(paste(italic(L)[Croot:Cveg])),
      limits = c(0, 2), 
      oob = scales::squish,
      midpoint = 1,
      reverse = TRUE
      ) +
    labs(
      title = display_titles[use_modl]
    ) +
    theme_void() +
    theme(
      legend.position = "none",
      plot.title = element_text(hjust = 0.5, size = 10) 
      )
}

plotlist <- purrr::map(
  model_order,
  ~ plot_by_model(., gdf)
)

legend_plot <- plotlist[[1]] +
    theme(legend.position = "right")

legend <- cowplot::get_legend(legend_plot)

map_grid <- cowplot::plot_grid(
  plotlist = plotlist,
  ncol = 3
)

final_plot <- cowplot::plot_grid(
  map_grid,
  legend,
  ncol = 2,
  rel_widths = c(1, 0.08)  # adjust the legend width as needed
)

final_plot

ggsave(
  filename = here("fig/map_dcroot_dcveg_bymodel.pdf"),
  height = 10, 
  width = 10
  )

ggsave(
  filename = here("fig/map_dcroot_dcveg_bymodel.png"),
  height = 10, 
  width = 10
  )
```

All models pooled.
```{r}
gdf_long_1x1deg <- gdf %>%
  select(-data) %>% 
  unnest(data_1x1deg) %>%
  group_by(lon, lat) %>% 
  summarise(
    l_croot_cveg_median = median(l_croot_cveg, na.rm = TRUE),
    l_croot_cveg_sd = sd(l_croot_cveg, na.rm = TRUE),
    nmodels = sum(!is.na(l_croot_cveg)),.groups = "drop"
  )%>%
  mutate(across(
    .cols = -c(lon, lat, nmodels), 
    .fns = ~ ifelse(nmodels < 8, NA, .x)
  ))
  
gg_map_agg <- ggplot() +
  geom_tile(
    data = gdf_long_1x1deg,
    aes(x = lon, y = lat, fill = l_croot_cveg_median),
    show.legend = TRUE
    ) +
  geom_sf(
    data = coast,
    colour = 'black',
    linewidth = 0.3
  )  +
  coord_sf(
    ylim = c(-60, 85),
    expand = FALSE
  ) +
  khroma::scale_fill_vik(
    name = expression(atop('Median of',italic(L)[Croot:Cveg])),
    limits = c(0.2, 1.8),  
    midpoint = 1,
    oob = scales::squish,
    reverse = TRUE
    ) +
  theme_void()


gg_map_aggSD <- ggplot() +
  geom_tile(
    data = gdf_long_1x1deg,
    aes(x = lon, y = lat, fill = l_croot_cveg_sd),
    show.legend = TRUE
    ) +
  geom_sf(
    data = coast,
    colour = 'black',
    linewidth = 0.3
  )  +
  coord_sf(
    ylim = c(-60, 85),
    expand = FALSE
  ) +
  khroma::scale_fill_grayC(
    name = expression(atop('SD of',italic(L)[Croot:Cveg])),
    limits = c(0, 2),  
    oob = scales::squish


    ) +
  theme_void()

gg_map_aggSD

```

#### Publication Fig. 5

```{r}
cowplot::plot_grid(
  gg3_l_croot_cveg,
  gg_map_agg,
  gg_map_aggSD,
  ncol = 1,
  labels = c("", "c", "d"),
  rel_heights = c(1, 0.9,0.9)
)

ggsave(filename = here("fig/fig5_combined_croot_cveg_sd.png"), height = 8, width = 6)
ggsave(filename = here("fig/fig5_combined_croot_cveg_sd.pdf"), height = 8, width = 6)
``` 

### Cwood vs Cveg

Taking the ratio of Cwood vs. Cveg here, both for their non-steady conditions (without applying Eq. 12 from paper). This is because we don't have NPP-wood for the correction. Assuming that changes in Cveg are dominated by changes in Cwood, the steady-state correction should cancel out anyways.

#### All models pooled

```{r}
set.seed(1982)

gdf_sampled <- gdf %>%
  mutate(data = purrr::map(data, ~slice_sample(., n = nsample, replace = TRUE))) %>%
  unnest(data) %>%
  dplyr::filter(
   !is.nan(dcwood), !is.nan(dcveg), 
    !is.infinite(dcwood), !is.infinite(dcveg),
    !is.nan(l_cwood_cveg), !is.infinite(l_cwood_cveg),l_cwood_cveg>-1000000,l_cwood_cveg<1000000)

# density scatter-plot for pooled data, balanced samples, original grid
gg_densityscatter <- gdf_sampled %>%
  ggplot(aes(x = dcveg, y = dcwood)) +
  geom_hex(bins = 50, show.legend = FALSE) +
  theme_classic() +
  geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 0, linetype = "dotted") +
  khroma::scale_fill_batlowW(trans = "log", reverse = TRUE) +
  xlim(-1, 2) + 
  ylim(-1, 2) +
  labs(
    x = expression(paste(Delta, italic(C)[veg], "/", italic(C)[veg])),
    y = expression(paste(Delta, italic(C)[root], "/", italic(C)[root])),
  )

# Number for text
print("Quartiles of L_cwood:Cveg")
print(format(quantile(gdf_sampled$l_cwood_cveg, probs =  c(0.25, 0.5, 0.75), na.rm = TRUE), digits = 3))

# Lables for Histogram  
labels <- gdf_sampled %>%
  ungroup()%>%
  mutate(l_larger = l_cwood_cveg > 1) %>%
  summarise(
    n_recs = n(),
    perc_larger = round(sum(l_larger, na.rm = TRUE)/n_recs,digits = 2),
    perc_smaller = 1 - perc_larger
    )

# Density plot
gg_density <- gdf_sampled %>% 
  ggplot(aes(x = l_cwood_cveg, y = after_stat(density))) +
  geom_density(fill = "grey70") +
  theme_classic() +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 1, linetype = "dotted") +
  xlim(-1, 2.5) + 
  geom_text(
    data = labels,
    aes(x = 0.5, y = Inf, label = perc_smaller),
    vjust= 1.5) +
  geom_text(
    data = labels,
    aes(x = 1.5, y = Inf, label = perc_larger),
    vjust= 1.5) +
  labs(
    x = expression(paste(italic(L)[Cwood:Cveg]))
  )

# Combine density scatter and density distribution plots
gg3_l_cwood_cveg <- cowplot::plot_grid(
  gg_densityscatter, 
  gg_density, 
  align = 'h',     # 'v' for vertical alignment
  axis = 'tb',
  ncol = 2,
  labels = c("a", "b")
)

ggsave(here("fig/cwood_cveg_pooled.pdf"), plot = gg3_l_cwood_cveg, width = 8, height = 4)
```


#### sensitivity test using JULES only

```{r cveg_star_croot_bymodl, fig.height=9, out.width="100%"}

set.seed(1982)

model_no <- which(gdf$modl=="JULES") 
model_df <- gdf$data[[model_no]]
robustbase::colMedians(as.matrix(model_df), na.rm = TRUE)
sum(!model_df$landCoverFrac==0)

gdf_sampled <- gdf %>%
  unnest(data) %>%
  filter(modl == "JULES")%>%
  mutate(l_cwood_star_cveg_star=dcwood_star/dcveg_star,
         l_cwood_cveg_star=dcwood/dcveg_star)%>%
  dplyr::filter(
    !is.nan(dcwood_star), !is.nan(dcveg_star), 
    !is.infinite(dcwood_star), !is.infinite(dcveg_star),
        !is.nan(dcwood), !is.nan(dcveg), 
    !is.infinite(dcwood), !is.infinite(dcveg),
    !is.nan(l_cwood_cveg), !is.infinite(l_cwood_cveg),
    !is.nan(l_cwood_star_cveg_star), !is.infinite(l_cwood_star_cveg_star),
    !is.nan(l_cwood_cveg_star), !is.infinite(l_cwood_cveg_star))

# density scatter

double_stara<-gdf_sampled %>%
  ggplot(aes(x = dcveg_star, y = dcwood_star)) +
  geom_hex(bins = 50, show.legend = FALSE) +
  theme_classic() +
  geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 0, linetype = "dotted") +
  khroma::scale_fill_batlowW(trans = "log", reverse = TRUE) +
  xlim(-1, 2) + 
  ylim(-1, 2) +
  labs(
    x = expression(paste(Delta, italic(C)[veg^"*"], "/", italic(C)[veg])),
    y = expression(paste(Delta, italic(C)[wood^"*"], "/", italic(C)[wood]))
  ) +
  facet_wrap(. ~ modl,  ncol = 3,  scales = "free") +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(size = 12, hjust = 0)
    )

one_stara<-gdf_sampled %>%
  ggplot(aes(x = dcveg_star, y = dcwood)) +
  geom_hex(bins = 50, show.legend = FALSE) +
  theme_classic() +
  geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 0, linetype = "dotted") +
  khroma::scale_fill_batlowW(trans = "log", reverse = TRUE) +
  xlim(-1, 2) + 
  ylim(-1, 2) +
  labs(
    x = expression(paste(Delta, italic(C)[veg^"*"], "/", italic(C)[veg])),
    y = expression(paste(Delta, italic(C)[wood], "/", italic(C)[wood]))
  ) +
  facet_wrap(. ~ modl,  ncol = 3,  scales = "free") +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(size = 12, hjust = 0)
    )


no_stara<-gdf_sampled %>%
  ggplot(aes(x = dcveg, y = dcwood)) +
  geom_hex(bins = 50, show.legend = FALSE) +
  theme_classic() +
  geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 0, linetype = "dotted") +
  khroma::scale_fill_batlowW(trans = "log", reverse = TRUE) +
  xlim(-1, 2) + 
  ylim(-1, 2) +
  labs(
    x = expression(paste(Delta, italic(C)[veg], "/", italic(C)[veg])),
    y = expression(paste(Delta, italic(C)[wood], "/", italic(C)[wood]))
  ) +
  facet_wrap(. ~ modl,  ncol = 3,  scales = "free") +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(size = 12, hjust = 0)
    )


# Lables for Histogram  
labels <- gdf_sampled %>%
  mutate(l_larger = l_cwood_star_cveg_star > 1) %>%
  group_by(modl) %>%
  summarise(
    n_recs = n(),
    perc_larger = round(sum(l_larger, na.rm = TRUE)/n_recs, digits = 2),
    perc_smaller = 1 - perc_larger
    )

# Histograms per model
double_star<-gdf_sampled %>%
  ggplot() +
  geom_density(aes(x = l_cwood_star_cveg_star, y = after_stat(density)), fill = "grey70") +
  theme_classic() +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 1, linetype = "dotted") +
  xlim(-1, 2.5) +
  geom_text(data = labels, aes(x = 0.5, y = Inf, label = perc_smaller), vjust =  1.5) +
  geom_text(data = labels, aes(x = 1.5, y = Inf, label = perc_larger), vjust = 1.5)+
  labs(
    x = expression(paste(italic(L)[Cwood^"*":Cveg^"*"]))
  ) +
  facet_wrap(. ~ modl, ncol = 3,  scales = "free") 

# Lables for Histogram  
labels <- gdf_sampled %>%
  mutate(l_larger = l_cwood_cveg_star > 1) %>%
  group_by(modl) %>%
  summarise(
    n_recs = n(),
    perc_larger = round(sum(l_larger, na.rm = TRUE)/n_recs, digits = 2),
    perc_smaller = 1 - perc_larger
    )

# Histograms per model
one_star<-gdf_sampled %>%
  ggplot() +
  geom_density(aes(x = l_cwood_cveg_star, y = after_stat(density)), fill = "grey70") +
  theme_classic() +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 1, linetype = "dotted") +
  xlim(-1, 2.5) +
  geom_text(data = labels, aes(x = 0.5, y = Inf, label = perc_smaller), vjust =  1.5) +
  geom_text(data = labels, aes(x = 1.5, y = Inf, label = perc_larger), vjust = 1.5)+
  labs(
    x = expression(paste(italic(L)[Cwood:Cveg^"*"]))
  ) +
  facet_wrap(. ~ modl, ncol = 3,  scales = "free") 


# Lables for Histogram  
labels <- gdf_sampled %>%
  mutate(l_larger = l_cwood_cveg > 1) %>%
  group_by(modl) %>%
  summarise(
    n_recs = n(),
    perc_larger = round(sum(l_larger, na.rm = TRUE)/n_recs, digits = 2),
    perc_smaller = 1 - perc_larger
    )

# Histograms per model
no_star<-gdf_sampled %>%
  ggplot() +
  geom_density(aes(x = l_cwood_cveg, y = after_stat(density)), fill = "grey70") +
  theme_classic() +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 1, linetype = "dotted") +
  xlim(-1, 2.5) +
  geom_text(data = labels, aes(x = 0.5, y = Inf, label = perc_smaller), vjust =  1.5) +
  geom_text(data = labels, aes(x = 1.5, y = Inf, label = perc_larger), vjust = 1.5)+
  labs(
    x = expression(paste(italic(L)[Cwood:Cveg]))
  ) +
  facet_wrap(. ~ modl, ncol = 3,  scales = "free") 
library(cowplot)
plot_grid(double_star,double_stara, one_star,one_stara, no_star,no_stara, ncol = 2,labels = "AUTO")
ggsave(filename = here('fig/sensitivity_test_star_cWood.png'),width=7,height=9)

```


#### By model

```{r cveg_cwood_bymodl, fig.height=9, out.width="100%"}
set.seed(1982)

gdf_sampled <- gdf %>%
  unnest(data) %>%
  bind_rows(
    .,
    gdf %>%
      mutate(n = purrr::map_int(data, ~nrow(.))) %>%
      mutate(data = purrr::map(data, ~slice_sample(., n = nsample))) %>% # smallest set here
      unnest(data) %>%
      mutate(modl = "ALL")%>%
      dplyr::filter(l_cwood_cveg>-1000000,l_cwood_cveg<1000000)
  ) %>%
  dplyr::filter(
    !is.nan(dcwood), !is.nan(dcveg), 
    !is.infinite(dcwood), !is.infinite(dcveg),
    !is.nan(l_cwood_cveg), !is.infinite(l_cwood_cveg))

gdf_sampled$modl <- factor(gdf_sampled$modl, levels = model_order_with_all)
# density scatter
gdf_sampled %>%
  ggplot(aes(x = dcveg, y = dcwood)) +
  geom_hex(bins = 50, show.legend = FALSE) +
  theme_classic() +
  geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
  geom_hline(yintercept = 0, linetype = "dotted") +
  # geom_vline(xintercept = 0, linetype = "dotted") +
  khroma::scale_fill_batlowW(trans = "log", reverse = TRUE) +
  xlim(-1, 2) + 
  ylim(-1, 2) +
  labs(
    x = expression(paste(Delta, italic(C)[veg], "/", italic(C)[veg])),
    y = expression(paste(Delta, italic(C)[wood], "/", italic(C)[wood]))
  ) +
  facet_wrap(. ~ modl,  ncol = 3,  scales = "free",labeller = as_labeller(display_titles_with_all)) +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(size = 12, hjust = 0)
    )

ggsave(here("fig/dcwood_dcveg_bymodl.pdf"), width = 9, height = 13)
ggsave(here("fig/dcwood_dcveg_bymodl.png"), width = 9, height = 13)

# we need to get some stat for the overall excel table
df_quantiles <- gdf_sampled %>%
  group_by(modl) %>%
  summarise(
    l_cwood_cveg_50 = quantile(l_cwood_cveg, probs = 0.5, na.rm = TRUE),
    l_cwood_cveg_iqr = IQR(l_cwood_cveg, na.rm = TRUE)
    ) 

summary_table <- left_join(summary_table, df_quantiles, by = 'modl')

# Lables for Histogram  
labels <- gdf_sampled %>%
  mutate(l_larger = l_cwood_cveg > 1) %>%
  group_by(modl) %>%
  summarise(
    n_recs = n(),
    perc_larger = round(sum(l_larger, na.rm = TRUE)/n_recs, digits = 2),
    perc_smaller = 1 - perc_larger
    )

# Histograms per model
gdf_sampled %>%
  ggplot() +
  geom_density(aes(x = l_cwood_cveg, y = after_stat(density)), fill = "grey70") +
  theme_classic() +
  geom_hline(yintercept = 0, linetype = "dotted") +
  # geom_vline(xintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 1, linetype = "dotted") +
  xlim(-1, 2.5) +
  geom_text(data = labels, aes(x = 0.5, y = Inf, label = perc_smaller), vjust =  1.5) +
  geom_text(data = labels, aes(x = 1.5, y = Inf, label = perc_larger), vjust = 1.5)+
  labs(
    x = expression(paste(italic(L)[Cwood:Cveg]))
  ) +
  facet_wrap(. ~ modl, ncol = 3,  scales = "free",labeller = as_labeller(display_titles_with_all)) 

ggsave(here("fig/hist_dcwood_dcveg_bymodl.pdf"), width = 9, height = 10.5)
ggsave(here("fig/hist_dcwood_dcveg_bymodl.png"), width = 9, height = 10.5)
```

#### Maps

By model.
```{r cveg_cwood_bymodl_map, fig.height=9, out.width="100%"}
plot_by_model <- function(use_modl, df){
  
  gdf_long <- df %>%
    unnest(data) %>%
    dplyr::filter(!is.nan(l_cwood_cveg), !is.infinite(l_cwood_cveg)) |> 
    filter(modl == use_modl)
  
  ggplot() +
    geom_tile(
      data = gdf_long,
      aes(x = lon, y = lat, fill = l_cwood_cveg),
      show.legend = TRUE
      ) +
    geom_sf(
      data = coast,
      colour = 'black',
      linewidth = 0.3
    )  +
    coord_sf(
      ylim = c(-60, 85),
      expand = FALSE
    ) +
    khroma::scale_fill_vik(
      name = expression(paste(italic(L)[Croot:Cveg])),
      limits = c(0, 2), 
      oob = scales::squish,
      midpoint = 1,
      reverse = TRUE
      ) +
    labs(
      title = display_titles[use_modl]
    ) +
    theme_void() +
    theme(
      legend.position = "none",
      plot.title = element_text(hjust = 0.5, size = 10) 
      )
}

plotlist <- purrr::map(
  model_order,
  ~ plot_by_model(., gdf)
)

legend_plot <- plotlist[[1]] +
    theme(legend.position = "right")

legend <- cowplot::get_legend(legend_plot)

map_grid <- cowplot::plot_grid(
  plotlist = plotlist,
  ncol = 3
)

final_plot <- cowplot::plot_grid(
  map_grid,
  legend,
  ncol = 2,
  rel_widths = c(1, 0.08)  # adjust the legend width as needed
)

final_plot

# # test
# plot_by_model("DLEM", gdf)

ggsave(
  filename = here("fig/map_dcroot_dcveg_bymodel.pdf"),
  height = 10, 
  width = 10
  )

ggsave(
  filename = here("fig/map_dcwood_dcveg_bymodel.png"),
  height = 10, 
  width = 10
  )
```

All models pooled.
```{r}
gdf_long_1x1deg <- gdf %>%
  select(-data) %>% 
  unnest(data_1x1deg) %>%
  group_by(lon, lat) %>% 
  summarise(
    l_cwood_cveg_median = median(l_cwood_cveg, na.rm = TRUE),
    l_cwood_cveg_sd = sd(l_cwood_cveg, na.rm = TRUE)
  )
  
gg_map_agg <- ggplot() +
  geom_tile(
    data = gdf_long_1x1deg,
    aes(x = lon, y = lat, fill = l_cwood_cveg_median),
    show.legend = TRUE
    ) +
  geom_sf(
    data = coast,
    colour = 'black',
    linewidth = 0.3
  )  +
  coord_sf(
    ylim = c(-60, 85),
    expand = FALSE
  ) +
  khroma::scale_fill_vik(
    name = expression(paste(italic(L)[Cwood:Cveg])),
    limits = c(0.2, 1.8),  
    midpoint = 1,
    oob = scales::squish,
    reverse = TRUE
    ) +
  theme_void()
```

#### SI Fig. XXX

```{r}
cowplot::plot_grid(
  gg3_l_cwood_cveg,
  gg_map_agg, 
  ncol = 1,
  labels = c("", "c"),
  rel_heights = c(1, 0.9)
)

ggsave(filename = here("fig/combined_cwood_cveg.png"), height = 5, width = 6)
ggsave(filename = here("fig/combined_cwood_cveg.pdf"), height = 5, width = 6)
``` 

### Croot vs Cveg*

Taking the ratio of Croot vs. Cveg where the latter is estimated for steady-state conditions. 

#### All models pooled

```{r}
set.seed(1982)

gdf_sampled <- gdf %>%
  mutate(data = purrr::map(data, ~slice_sample(., n = nsample))) %>%
  unnest(data) %>%
  dplyr::filter(
    !is.nan(dcroot), !is.nan(dcveg_star), 
    !is.infinite(dcroot), !is.infinite(dcveg_star),
    !is.nan(l_croot_cveg_star), !is.infinite(l_croot_cveg_star))

# density scatter-plot for pooled data, balanced samples, original grid
gg_densityscatter <- gdf_sampled %>%
  ggplot(aes(x = dcveg_star, y = dcroot)) +
  geom_hex(bins = 50, show.legend = FALSE) +
  theme_classic() +
  geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 0, linetype = "dotted") +
  khroma::scale_fill_batlowW(trans = "log", reverse = TRUE) +
  xlim(-1, 2) + 
  ylim(-1, 2) +
  labs(
    x = expression(paste(Delta, italic(C)[veg^"*"], "/", italic(C)[veg])),
    y = expression(paste(Delta, italic(C)[root], "/", italic(C)[root])),
  )

# Number for text
print("Quartiles of L_Croot:Cveg*")
print(format(quantile(gdf_sampled$l_croot_cveg_star, probs =  c(0.25, 0.5, 0.75), na.rm = TRUE), digits = 3))

# Lables for Histogram  
labels <- gdf_sampled %>%
  ungroup()%>%
  mutate(l_larger = l_croot_cveg_star > 1) %>%
  summarise(
    n_recs = n(),
    perc_larger = round(sum(l_larger, na.rm = TRUE)/n_recs, digits = 2),
    perc_smaller = 1 - perc_larger
    )

# Density plot
gg_density <- gdf_sampled %>% 
  ggplot(aes(x = l_croot_cveg_star, y = after_stat(density))) +
  geom_density(fill = "grey70") +
  theme_classic() +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 1, linetype = "dotted") +
  xlim(-1, 2.5) + 
  geom_text(
    data = labels,
    aes(x = 0.5, y = Inf, label = perc_smaller),
    vjust= 1.5) +
  geom_text(
    data = labels,
    aes(x = 1.5, y = Inf, label = perc_larger),
    vjust= 1.5) +
  labs(
    x = expression(paste(italic(L)[Croot:Cveg^"*"]))
  )

# Combine density scatter and density distribution plots
gg3_l_croot_cveg_star <- cowplot::plot_grid(
  gg_densityscatter, 
  gg_density, 
  align = 'h',     # 'v' for vertical alignment
  axis = 'tb',
  ncol = 2,
  labels = c("a", "b")
)

ggsave(here("fig/croot_cveg_star_pooled.pdf"), plot = gg3_l_croot_cveg_star, width = 8, height = 4)
```

#### By model

```{r cveg_star_croot_bymodl, fig.height=9, out.width="100%"}
set.seed(1982)

gdf_sampled <- gdf %>%
  unnest(data) %>%
  bind_rows(
    .,
    gdf %>%
      mutate(n = purrr::map_int(data, ~nrow(.))) %>%
      mutate(data = purrr::map(data, ~slice_sample(., n = nsample, replace = TRUE))) %>% # smallest set here
      unnest(data) %>%
      mutate(modl = "ALL")
  ) %>%
  dplyr::filter(!is.nan(dcroot), !is.nan(dcveg_star), !is.infinite(dcroot), !is.infinite(dcveg_star))
gdf_sampled$modl <- factor(gdf_sampled$modl, levels = model_order_with_all)
# density scatter
gdf_sampled %>%
  ggplot(aes(x = dcveg_star, y = dcroot)) +
  geom_hex(bins = 50, show.legend = FALSE) +
  theme_classic() +
  geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
  geom_hline(yintercept = 0, linetype = "dotted") +
  # geom_vline(xintercept = 0, linetype = "dotted") +
  khroma::scale_fill_batlowW(trans = "log", reverse = TRUE) +
  xlim(-1, 2) + 
  ylim(-1, 2) +
  labs(
    x = expression(paste(Delta, italic(C)[veg^"*"], "/", italic(C)[veg])),
    y = expression(paste(Delta, italic(C)[root], "/", italic(C)[root]))
  ) +
  facet_wrap(. ~ modl,  ncol = 3,  scales = "free",labeller = as_labeller(display_titles_with_all)) +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(size = 12, hjust = 0)
    )

ggsave(here("fig/dcroot_dcveg_star_bymodl.pdf"), width = 9, height = 13)
ggsave(here("fig/dcroot_dcveg_star_bymodl.png"), width = 9, height = 13)

# we need to get some stat for the overall excel table
df_quantiles <- gdf_sampled %>%
  group_by(modl) %>%
  summarise(
    l_croot_cveg_star_50 = quantile(l_croot_cveg_star, probs = 0.5, na.rm = TRUE),
    l_croot_cveg_star_iqr = IQR(l_croot_cveg_star, na.rm = TRUE)
    ) 

summary_table <- left_join(summary_table, df_quantiles, by = 'modl')

# Lables for Histogram  
labels <- gdf_sampled %>%
  mutate(l_larger = l_croot_cveg_star > 1) %>%
  group_by(modl) %>%
  summarise(
    n_recs = n(),
    perc_larger = round(sum(l_larger, na.rm = TRUE)/n_recs, digits = 2),
    perc_smaller = 1 - perc_larger
    )

# Histograms per model
gdf_sampled %>%
  ggplot() +
  geom_density(aes(x = l_croot_cveg_star, y = after_stat(density)), fill = "grey70") +
  theme_classic() +
  geom_hline(yintercept = 0, linetype = "dotted") +
  # geom_vline(xintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 1, linetype = "dotted") +
  xlim(-1, 2.5) +
  geom_text(data = labels, aes(x = 0.5, y = Inf, label = perc_smaller), vjust =  1.5) +
  geom_text(data = labels, aes(x = 1.5, y = Inf, label = perc_larger), vjust = 1.5)+
  labs(
    x = expression(paste(italic(L)[Croot:Cveg^"*"]))
  ) +
  facet_wrap(. ~ modl, ncol = 3,  scales = "free",labeller = as_labeller(display_titles_with_all)) 

ggsave(here("fig/hist_dcroot_dcveg_star_bymodl.pdf"), width = 9, height = 10.5)
ggsave(here("fig/hist_dcroot_dcveg_star_bymodl.png"), width = 9, height = 10.5)
```



#### Maps

By model.
```{r cveg_star_croot_bymodl_maps, fig.height=9, out.width="100%"}
plot_by_model <- function(use_modl, df){
  
  gdf_long <- df %>%
    unnest(data) %>%
    dplyr::filter(!is.nan(l_croot_cveg_star), !is.infinite(l_croot_cveg_star)) |> 
    filter(modl == use_modl)
  
  ggplot() +
    geom_tile(
      data = gdf_long,
      aes(x = lon, y = lat, fill = l_croot_cveg_star),
      show.legend = TRUE
      ) +
    geom_sf(
      data = coast,
      colour = 'black',
      linewidth = 0.3
    )  +
    coord_sf(
      ylim = c(-60, 85),
      expand = FALSE
    ) +
    khroma::scale_fill_vik(
      name = expression(paste(italic(L)[Croot:Cveg^"*"])),
      limits = c(0, 2), 
      oob = scales::squish,
      midpoint = 1,
      reverse = TRUE
      ) +
    labs(
      title = display_titles[use_modl]
    ) +
    theme_void() +
    theme(
      legend.position = "none",
      plot.title = element_text(hjust = 0.5, size = 10) 
      )
}

plotlist <- purrr::map(
  model_order,
  ~ plot_by_model(., gdf)
)

legend_plot <- plotlist[[1]] +
    theme(legend.position = "right")

legend <- cowplot::get_legend(legend_plot)

map_grid <- cowplot::plot_grid(
  plotlist = plotlist,
  ncol = 3
)

final_plot <- cowplot::plot_grid(
  map_grid,
  legend,
  ncol = 2,
  rel_widths = c(1, 0.08)  # adjust the legend width as needed
)

final_plot

ggsave(
  filename = here("fig/map_dcroot_dcveg_star_bymodel.pdf"),
  height = 10, 
  width = 10
  )

ggsave(
  filename = here("fig/map_dcroot_dcveg_star_bymodel.png"),
  height = 10, 
  width = 10
  )
```

All models pooled.
```{r}
gdf_long_1x1deg <- gdf %>%
  select(-data) %>% 
  unnest(data_1x1deg) %>%
  group_by(lon, lat) %>% 
  summarise(
    l_croot_cveg_star_median = median(l_croot_cveg_star, na.rm = TRUE),
    l_croot_cveg_star_sd = sd(l_croot_cveg_star, na.rm = TRUE)
  )
  
gg_map_agg <- ggplot() +
  geom_tile(
    data = gdf_long_1x1deg,
    aes(x = lon, y = lat, fill = l_croot_cveg_star_median),
    show.legend = TRUE
    ) +
  geom_sf(
    data = coast,
    colour = 'black',
    linewidth = 0.3
  )  +
  coord_sf(
    ylim = c(-60, 85),
    expand = FALSE
  ) +
  khroma::scale_fill_vik(
    name = expression(paste(italic(L)[Croot:Cveg^"*"])),
    limits = c(0.2, 1.8),  
    midpoint = 1,
    oob = scales::squish,
    reverse = TRUE
    ) +
  theme_void()
```

### Cleaf vs Cveg*

#### All models pooled

```{r}
set.seed(1982)

gdf_sampled <- gdf %>%
  mutate(data = purrr::map(data, ~slice_sample(., n = nsample))) %>%
  unnest(data) %>%
  dplyr::filter(
    !is.nan(dcleaf), !is.nan(dcveg_star), 
    !is.infinite(dcleaf), !is.infinite(dcveg_star),
    !is.nan(l_cleaf_cveg_star), !is.infinite(l_cleaf_cveg_star))

# density scatter-plot for pooled data, balanced samples, original grid
gg_densityscatter <- gdf_sampled %>%
  ggplot(aes(x = dcveg_star, y = dcleaf)) +
  geom_hex(bins = 50, show.legend = FALSE) +
  theme_classic() +
  geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 0, linetype = "dotted") +
  khroma::scale_fill_batlowW(trans = "log", reverse = TRUE) +
  xlim(-1, 2) + 
  ylim(-1, 2) +
  labs(
    x = expression(paste(Delta, italic(C)[veg^"*"], "/", italic(C)[veg])),
    y = expression(paste(Delta, italic(C)[leaf], "/", italic(C)[leaf])),
  )

# Number for text
print("Quartiles of L_Cleaf:Cveg*")
print(format(quantile(gdf_sampled$l_cleaf_cveg_star, probs =  c(0.25, 0.5, 0.75), na.rm = TRUE), digits = 3))

# Lables for Histogram  
labels <- gdf_sampled %>%
  ungroup()%>%
  mutate(l_larger = l_cleaf_cveg_star > 1) %>%
  summarise(
    n_recs = n(),
    perc_larger = round(sum(l_larger, na.rm = TRUE)/n_recs, digits = 2),
    perc_smaller = 1 - perc_larger
    )

# Density plot
gg_density <- gdf_sampled %>% 
  ggplot(aes(x = l_cleaf_cveg_star, y = after_stat(density))) +
  geom_density(fill = "grey70") +
  theme_classic() +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 1, linetype = "dotted") +
  xlim(-1, 2.5) + 
  geom_text(
    data = labels,
    aes(x = 0.5, y = Inf, label = perc_smaller),
    vjust= 1.5) +
  geom_text(
    data = labels,
    aes(x = 1.5, y = Inf, label = perc_larger),
    vjust= 1.5) +
  labs(
    x = expression(paste(italic(L)[Cleaf:Cveg^"*"]))
  )

# Combine density scatter and density distribution plots
gg3_l_cleaf_cveg_star <- cowplot::plot_grid(
  gg_densityscatter, 
  gg_density, 
  align = 'h',     # 'v' for vertical alignment
  axis = 'tb',
  ncol = 2,
  labels = c("a", "b")
)

ggsave(here("fig/cleaf_cveg_star_pooled.pdf"), plot = gg3_l_cleaf_cveg_star, width = 8, height = 4)
```

#### By model

```{r cveg_star_cleaf_bymodl, fig.height=9, out.width="100%"}
set.seed(1982)

gdf_sampled <- gdf %>%
  unnest(data) %>%
  bind_rows(
    .,
    gdf %>%
      mutate(n = purrr::map_int(data, ~nrow(.))) %>%
      mutate(data = purrr::map(data, ~slice_sample(., n = nsample, replace = TRUE))) %>% # smallest set here
      unnest(data) %>%
      mutate(modl = "ALL")
  ) %>%
  dplyr::filter(
    !is.nan(dcleaf), !is.nan(dcveg_star), 
    !is.infinite(dcleaf), !is.infinite(dcveg_star))
gdf_sampled$modl <- factor(gdf_sampled$modl, levels = model_order_with_all)

# density scatter
gdf_sampled %>%
  ggplot(aes(x = dcveg_star, y = dcleaf)) +
  geom_hex(bins = 50, show.legend = FALSE) +
  theme_classic() +
  geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
  geom_hline(yintercept = 0, linetype = "dotted") +
  # geom_vline(xintercept = 0, linetype = "dotted") +
  khroma::scale_fill_batlowW(trans = "log", reverse = TRUE) +
  xlim(-1, 2) + 
  ylim(-1, 2) +
  labs(
    x = expression(paste(Delta, italic(C)[veg^"*"], "/", italic(C)[veg])),
    y = expression(paste(Delta, italic(C)[leaf], "/", italic(C)[leaf]))
  ) +
  facet_wrap(. ~ modl,  ncol = 3,  scales = "free",labeller = as_labeller(display_titles_with_all)) +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(size = 12, hjust = 0)
    )

ggsave(here("fig/dcleaf_dcveg_star_bymodl.pdf"), width = 9, height = 13)
ggsave(here("fig/dcleaf_dcveg_star_bymodl.png"), width = 9, height = 13)

# we need to get some stat for the overall excel table
df_quantiles <- gdf_sampled %>%
  group_by(modl) %>%
  summarise(
    l_cleaf_cveg_star_50 = quantile(l_cleaf_cveg_star, probs = 0.5, na.rm = TRUE),
    l_cleaf_cveg_star_iqr = IQR(l_cleaf_cveg_star, na.rm = TRUE)
    ) 

summary_table <- left_join(summary_table, df_quantiles, by = 'modl')

# Lables for Histogram  
labels <- gdf_sampled %>%
  mutate(l_larger = l_cleaf_cveg_star > 1) %>%
  group_by(modl) %>%
  summarise(
    n_recs = n(),
    perc_larger = round(sum(l_larger, na.rm = TRUE)/n_recs, digits = 2),
    perc_smaller = 1 - perc_larger
    )

# Histograms per model
gdf_sampled %>%
  ggplot() +
  geom_density(aes(x = l_cleaf_cveg_star, y = after_stat(density)), fill = "grey70") +
  theme_classic() +
  geom_hline(yintercept = 0, linetype = "dotted") +
  # geom_vline(xintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 1, linetype = "dotted") +
  xlim(-1, 2.5) +
  geom_text(data = labels, aes(x = 0.5, y = Inf, label = perc_smaller), vjust =  1.5) +
  geom_text(data = labels, aes(x = 1.5, y = Inf, label = perc_larger), vjust = 1.5)+
  labs(
    x = expression(paste(italic(L)[Cleaf:Cveg^"*"]))
  ) +
  facet_wrap(. ~ modl, ncol = 3,  scales = "free",labeller = as_labeller(display_titles_with_all)) 

ggsave(here("fig/hist_dcleaf_dcveg_star_bymodl.pdf"), width = 9, height = 10.5)
ggsave(here("fig/hist_dcleaf_dcveg_star_bymodl.png"), width = 9, height = 10.5)
```

#### Maps

By model.
```{r cveg_star_cleaf_bymodl_maps, fig.height=9, out.width="100%"}
plot_by_model <- function(use_modl, df){
  
  gdf_long <- df %>%
    unnest(data) %>%
    dplyr::filter(!is.nan(l_cleaf_cveg), !is.infinite(l_cleaf_cveg)) |> 
    filter(modl == use_modl)
  
  ggplot() +
    geom_tile(
      data = gdf_long,
      aes(x = lon, y = lat, fill = l_cleaf_cveg_star),
      show.legend = TRUE
      ) +
    geom_sf(
      data = coast,
      colour = 'black',
      linewidth = 0.3
    )  +
    coord_sf(
      ylim = c(-60, 85),
      expand = FALSE
    ) +
    khroma::scale_fill_vik(
      name = expression(paste(italic(L)[Cleaf:Cveg^"*"])),
      limits = c(0, 2), 
      oob = scales::squish,
      midpoint = 1,
      reverse = TRUE
      ) +
    labs(
     title = display_titles[use_modl]
    ) +
    theme_void() +
    theme(
      legend.position = "none",
      plot.title = element_text(hjust = 0.5, size = 10) 
      )
}

plotlist <- purrr::map(
   model_order,
  ~ plot_by_model(., gdf)
)

legend_plot <- plotlist[[1]] +
    theme(legend.position = "right")

legend <- cowplot::get_legend(legend_plot)

map_grid <- cowplot::plot_grid(
  plotlist = plotlist,
  ncol = 3
)

final_plot <- cowplot::plot_grid(
  map_grid,
  legend,
  ncol = 2,
  rel_widths = c(1, 0.08)  # adjust the legend width as needed
)

final_plot

ggsave(
  filename = here("fig/map_dcleaf_dcveg_star_bymodel.pdf"),
  height = 10, 
  width = 10
  )

ggsave(
  filename = here("fig/map_dcleaf_dcveg_star_bymodel.png"),
  height = 10, 
  width = 10
  )
```

All models pooled.
```{r}
gdf_long_1x1deg <- gdf %>%
  select(-data) %>% 
  unnest(data_1x1deg) %>%
  group_by(lon, lat) %>% 
  summarise(
    l_cleaf_cveg_star_median = median(l_cleaf_cveg_star, na.rm = TRUE),
    l_cleaf_cveg_star_sd = sd(l_cleaf_cveg_star, na.rm = TRUE)
  )
  
gg_map_agg <- ggplot() +
  geom_tile(
    data = gdf_long_1x1deg,
    aes(x = lon, y = lat, fill = l_cleaf_cveg_star_median),
    show.legend = TRUE
    ) +
  geom_sf(
    data = coast,
    colour = 'black',
    linewidth = 0.3
  )  +
  coord_sf(
    ylim = c(-60, 85),
    expand = FALSE
  ) +
  khroma::scale_fill_vik(
    name = expression(paste(italic(L)[Cleaf:Cveg^"*"])),
    limits = c(0.2, 1.8),  
    midpoint = 1,
    oob = scales::squish,
    reverse = TRUE
    ) +
  theme_void()
```

###  cRoot ratio

#### Publication Fig. supplementary

```{r}
havealook <- gdf %>% #Prepare the dataset, this is similar to the histogram above
  unnest(data)  %>%
  filter(!modl %in% c('LPJ-GUESS','JSBACH','DLEM','LPJ','VISIT'))%>%
  dplyr::filter(!is.nan(croot_init), !is.nan(cveg_init), !is.infinite(croot_init), !is.infinite(cveg_init)) %>%
  mutate(root_shoot = croot_init/cveg_init )%>%
  filter(root_shoot<100000, root_shoot>-1000000) %>% # to remove NA and infinite
  mutate(lon=ifelse(lon>180,lon-360,lon)) #some maps have longitude from 0 to 360, need to convert to -180 to 180

bymodel_map <- ggplot() +
  geom_raster(
    data = havealook,
    aes(x = lon, y = lat, fill = root_shoot),
    show.legend = TRUE
    ) +
  geom_sf(
    data = coast,
    colour = 'black',
    linewidth = 0.3
  )  +
  coord_sf(
    ylim = c(-60, 85),
    expand = FALSE
  ) +
  khroma::scale_fill_vik(
    name = "",
    limits = c(0,1), 
    midpoint = 0.5
    ) +
  theme_void() +
  labs(
      title = expression(paste('[Croot/Cveg]','  as median across models')) 
  ) +
  facet_wrap(. ~ modl, ncol = 3)

ggsave(bymodel_map,filename = here("fig/rootshootration_trendyv11_bymodelv1.pdf"),height=10, width = 10)

havealook <-havealook%>%
    mutate(root_shoot=ifelse(root_shoot>0.1,0.1,root_shoot),
         root_shoot=ifelse(root_shoot<0,0,root_shoot))

bymodel_map <- ggplot() +
  geom_raster(
    data = havealook,
    aes(x = lon, y = lat, fill = root_shoot),
    show.legend = TRUE
    ) +
  geom_sf(
    data = coast,
    colour = 'black',
    linewidth = 0.3
  )  +
  coord_sf(
    ylim = c(-60, 85),
    expand = FALSE
  ) +
  khroma::scale_fill_vik(
    name = "",
    limits = c(0,0.1), 
    midpoint = 0.05
    ) +
  theme_void() +
  labs(
      title = expression(paste('[Croot/Cveg]')) 
  ) +
  facet_wrap(. ~ modl, ncol = 3)

ggsave(bymodel_map,filename = here("fig/rootshootration_trendyv11_bymodelv2.png"),height=10, width = 10)

```

###  cVeg versus cVeg_star


```{r}



havealook <- gdf %>% #Prepare the dataset, this is similar to the histogram above
  unnest(data)  %>%
  dplyr::filter(!is.nan(dcveg), !is.nan(dcveg_star), !is.infinite(dcveg), !is.infinite(dcveg_star))%>%
  mutate(star_effect = dcveg/dcveg_star )%>%
  filter(star_effect<100000, star_effect>-1000000) %>% # to remove NA and infinite
  mutate(lon=ifelse(lon>180,lon-360,lon)) #some maps have longitude from 0 to 360, need to convert to -180 to 180

bymodel_map1 <- ggplot() +
  geom_raster(
    data = havealook,
    aes(x = lon, y = lat, fill = star_effect),
    show.legend = TRUE
    ) +
  geom_sf(
    data = coast,
    colour = 'black',
    linewidth = 0.3
  )  +
  coord_sf(
    ylim = c(-60, 85),
    expand = FALSE
  ) +
  khroma::scale_fill_vik(
    name = "",
    limits = c(0,2), 
    midpoint = 1
    ) +
  theme_void() +
  labs(
    title =  expression(italic(R)["Cveg"] / italic(R)["Cveg*"])
  )+
  facet_wrap(. ~ modl, ncol = 3)

ggsave(bymodel_map1,filename = here("fig/star_effect_trendyv11_bymodelv1.pdf"),height=10, width = 10)

boxplot_stareffect <- ggplot(havealook, aes(y = star_effect, x = modl)) +
  geom_boxplot() +
  coord_flip() +
  scale_y_continuous(limits = c(0.5, 1.4)) +
  labs(
    x = expression(italic(R)["Cveg"] / italic(R)["Cveg*"]),
    y = NULL
  )
  
ggsave(boxplot_stareffect,filename = here("fig/star_effect_boxPlottrendyv11_bymodelv1.png"),height=5, width = 6)

bymodel_map <- ggplot() +
  geom_raster(
    data = havealook,
    aes(x = lon, y = lat, fill = dcveg),
    show.legend = TRUE
    ) +
  geom_sf(
    data = coast,
    colour = 'black',
    linewidth = 0.3
  )  +
  coord_sf(
    ylim = c(-60, 85),
    expand = FALSE
  ) +
  khroma::scale_fill_vik(
    name = "",
    limits = c(-1,1), 
    midpoint = 0
    ) +
  theme_void() +
  labs(
      title = expression(paste('[Croot/Cveg]')) 
  ) +
  facet_wrap(. ~ modl, ncol = 3)

ggsave(bymodel_map,filename = here("fig/dcveg_trendyv11_bymodelv2.png"),height=10, width = 10)
```

###  change in autotrophic respiration

#### Publication Fig. supplementary

```{r}


havealook <- gdf %>% #Prepare the dataset, this is similar to the histogram above
  unnest(data)  %>%
  mutate(Ra_final = gpp_final-npp_final ,
         Ra_init=gpp_init-npp_init,
         dRa = (Ra_final-Ra_init)/Ra_init)%>%
  dplyr::filter(!is.nan(dRa), !is.nan(dRa), !is.infinite(dRa), !is.infinite(dRa))%>%
   # to remove NA and infinite
  mutate(lon=ifelse(lon>180,lon-360,lon)) #some maps have longitude from 0 to 360, need to convert to -180 to 180


bymodel_map1 <- ggplot() +
  geom_raster(
    data = havealook,
    aes(x = lon, y = lat, fill = dRa),
    show.legend = TRUE
    ) +
  geom_sf(
    data = coast,
    colour = 'black',
    linewidth = 0.3
  )  +
  coord_sf(
    ylim = c(-60, 85),
    expand = FALSE
  ) +
  khroma::scale_fill_vik(
    name = "",
    limits = c(-0.5,0.5), 
    midpoint = 0
    ) +
  theme_void() +
  labs(
    title =  expression(italic(R)["Ra"] )
  )+
  facet_wrap(. ~ modl, ncol = 3)


ggsave(bymodel_map1,filename = here("fig/RRa_trendyv11_bymodelv1.png"),height=10, width = 10)

```

## Summary table (Publication Fig. 2)

```{r summary_table, fig.height=9, out.width="80%"}

#write.csv(summary_table,file = here("data/summary_table_figure2.csv"))

summary_table <- read.csv(here("data/summary_table_figure2.csv"))
cn_id <- tibble(
  cn_coupled = c(TRUE, FALSE),
  cn_label = c("C-N", "    C-only")
)

veg_demo_id<- tibble(
  veg_demo = c(TRUE, FALSE),
  cn_label = c("Veg. Demo.", "no Veg. Demo.")
)

model_order <- c(
  "VISIT-NIES", "CLASSIC", "IBIS", "LPJ", "VISIT", 
  "ISBA-CTRIP", "ORCHIDEE", "CLM5.0", "DLEM", "ISAM", 
  "JSBACH", "JULES", "LPX-Bern", "SDGVM", "CABLE-POP", "LPJ-GUESS"
)


df <- summary_table %>% 
  filter(modl != "ALL") %>%
  mutate(modl = factor(modl, levels = rev(model_order)))

column_order <- c(
  "l_npp_gpp_50", "l_cveg_star_npp_50", "l_croot_cveg_50",
  "l_cwood_cveg_50", "l_cleaf_cveg_star_50", "cn_coupled", "veg_demo"
)

column_labels <- c(
  expression(paste(italic(L)[NPP:GPP])),
  expression(paste(italic(L)[Cveg^"*":NPP])),
  expression(paste(italic(L)[Croot:Cveg])),
  expression(paste(italic(L)[Cwood:Cveg])),
  expression(paste(italic(L)[Cleaf:Cveg^"*"])),
  "C-N",
  "Veg. Demo."
)

df_long <- df %>%
  select(modl, all_of(column_order[1:5])) %>%
  pivot_longer(-modl, names_to = "link", values_to = "value") %>%
  mutate(
    link = factor(link, levels = column_order),
    row_type = "value",
    label = sprintf("%.2f", value)
  )

df_notes <- df %>%
  select(modl, cn_coupled, veg_demo) %>%
  pivot_longer(-modl, names_to = "link", values_to = "note") %>%
  mutate(
    link = factor(link, levels = column_order),
    row_type = "note",
    value = NA_real_,
    label = ifelse(note, "Yes", "No")
  )


df_plot <- bind_rows(df_long, df_notes)

gg1 <- ggplot(df_plot, aes(x = link, y = modl, fill = value)) +
  geom_tile(color = "white", show.legend = FALSE) +
  geom_text(
    data = filter(df_plot, row_type == "value"),
    aes(label = label),
    color = "black"
  ) +
  geom_text(
    data = filter(df_plot, row_type == "note"),
    aes(label = label),
    color = "black"
  ) +
   khroma::scale_fill_vik(
    name = expression(paste(italic(L))),
    limits = c(0.2, 1.8),
    midpoint = 1,
    reverse = TRUE,
    na.value = "white"
  ) +
  scale_x_discrete(labels = column_labels, position = "top") +
  scale_y_discrete() + 
  coord_fixed(ratio = 0.5) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.grid = element_blank(),
    legend.position = "bottom"
  ) +
  labs(fill = "Value", y = "", x = "")

# C-N vs C-only

df_long <- df %>%
  group_by(cn_coupled) |> 
  summarise(across(all_of(column_order[1:5]), ~ median(.x, na.rm = TRUE))) |>
  pivot_longer(-cn_coupled, names_to = "link", values_to = "value")

df_long <- df_long |> 
  bind_rows(
    expand_grid(
      cn_coupled = unique(df_long$cn_coupled), 
      link = c("Extra1", "Extra2"), 
      value = as.numeric(NA)
    )
  ) |> 
  mutate(link = factor(link, levels = c(column_order[1:5], "Extra1", "Extra2"))) |> 
  left_join(cn_id, by = "cn_coupled")


gg2 <- ggplot(df_long, aes(x = link, y = cn_label, fill = value)) +
  geom_tile(color = "white", show.legend = FALSE) +
  geom_text(aes(label = round(value, 2)), color = "black") +
  khroma::scale_fill_vik(
    name = expression(paste(italic(L))),
    limits = c(0.2, 1.8),  
    oob = scales::squish,
    midpoint = 1,
    reverse = TRUE,
    na.value = "white" # This colors the new NA tiles grey
  ) +
  scale_x_discrete(labels = c(column_labels[1:5], "New Col 1", "New Col 2"), position = "top") +
  coord_fixed(ratio = 0.5) +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    panel.grid = element_blank(),
    legend.position = "bottom"
  ) +
  labs(fill = "Value", y = "", x = "")

# veg demo or not
df_long <- df %>%
  group_by(veg_demo) |> 
  summarise(across(all_of(column_order[1:5]), ~ median(.x, na.rm = TRUE))) |> 
  pivot_longer(-veg_demo, names_to = "link", values_to = "value")

new_cols <- crossing(
  veg_demo = unique(df_long$veg_demo), 
  link = c("Extra_1", "Extra_2")
) |> 
  mutate(value = as.numeric(NA))

df_long <- df_long |> 
  bind_rows(new_cols) |> 
  mutate(link = factor(link, levels = c(column_order, "Extra_1", "Extra_2"))) |> 
  left_join(veg_demo_id, by = "veg_demo")

gg2veg <- ggplot(df_long, aes(x = link, y = cn_label, fill = value)) +
  geom_tile(color = "white", show.legend = FALSE) +
  geom_text(aes(label = round(value, 2)), color = "black") +
  khroma::scale_fill_vik(
    name = expression(paste(italic(L))),
    limits = c(0.2, 1.8),  
    oob = scales::squish,
    midpoint = 1,
    reverse = TRUE,
    na.value = "white"
  ) +
  # Add labels for the two new columns here
  scale_x_discrete(labels = c(column_labels, "New 1", "New 2"), position = "top") +
  coord_fixed(ratio = 0.5) +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    panel.grid = element_blank(),
    legend.position = "bottom"
  ) +
  labs(fill = "Value", y = "", x = "")

# aggregated across all models
df_long <- df %>%
  summarise(across(all_of(column_order[1:5]), ~ median(.x, na.rm = TRUE))) |> 
  pivot_longer(cols = everything(), names_to = "link", values_to = "value")


new_cols <- tibble(
  link = c("Extra_1", "Extra_2"), 
  value = as.numeric(NA)
)

df_long <- df_long |> 
  bind_rows(new_cols) |> 
  mutate(link = factor(link, levels = c(column_order, "Extra_1", "Extra_2"))) |> 
  mutate(label = "ALL")

# 3. Plotting (Your existing ggplot code remains the same)
gg3 <- ggplot(df_long, aes(x = link, y = label, fill = value)) +
  geom_tile(color = "white", show.legend = TRUE) +
  geom_text(aes(label = round(value, 2)), color = "grey30") +
  khroma::scale_fill_vik(
    name = expression(paste(italic(L))),
    limits = c(0.2, 1.8),  
    oob = scales::squish,
    midpoint = 1,
    reverse = TRUE,
    na.value = "white" # This will color your new NA columns grey
  ) +
  scale_x_discrete(labels = c(column_labels, "New Col 1", "New Col 2"), position = "top") +
  coord_fixed(ratio = 0.5) +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    panel.grid = element_blank(),
    legend.position = "none"
  ) +
  labs(fill = "Value", y = "", x = "")

df_long_gg4 <- df_long

df_long_gg4$label <- "Observation"
df_long_gg4$value <- c(NA, 0.63, 1.22, 1.09, NA, NA, NA)
# link column remains unchanged

# Plot gg4 (same structure as gg3)
gg4 <- ggplot(df_long_gg4, aes(x = link, y = label, fill = value)) +
  geom_tile(color = "white", show.legend = TRUE) +
  geom_text(aes(label = round(value, 2)), color = "grey30") +
  khroma::scale_fill_vik(
    name = expression(paste(italic(L))),
    limits = c(0.2, 1.8),  
    oob = scales::squish,
    midpoint = 1,
    reverse = TRUE,
    na.value = "white"
  ) +
  scale_x_discrete(labels = c(column_labels, "New Col 1", "New Col 2"), position = "top") +
  coord_fixed(ratio = 0.5) +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    panel.grid = element_blank(),
    legend.position = "bottom"
  ) +
  labs(fill = "Value", y = "", x = "")


# get legend
# Dummy plot just to extract the legend
legend_plot <- ggplot(df_long, aes(x = link, y = label, fill = value)) +
  geom_tile() +
  khroma::scale_fill_vik(
    name = expression(paste(italic(L))),
    limits = c(0.2, 1.8),  
    oob = scales::squish,
    midpoint = 1,
    reverse = TRUE,
    na.value = "grey70"
    ) +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal"
    )

# Extract legend object
legend <- cowplot::get_legend(legend_plot)

# heatmaps <- cowplot::plot_grid(
#   gg1,
#   gg_yesno,
#   gg2, 
#   NULL,
#   gg2veg,
#    NULL,
#   gg3,
#   NULL,
#   legend,
#   ncol = 2,
#   align = "hv",    # vertical alignment
#   axis = "lr",    # align left and right margins (y-axis)
#   rel_heights = c(1, 0.182, 0.182, 0.127, 0.127),
#   labels = letters[1:3]
# )

# install.packages("egg")
library(egg)
heatmaps <- egg::ggarrange(
  gg1, gg2, gg2veg, gg3, gg4,
  ncol = 1,
  heights = c(1, 0.189, 0.189,0.189, 0.25), # Adjust for 4 plots
  labels = c("a", "b", "c", "d","e"),
  debug = FALSE # Set to TRUE to see the alignment boxes
)

ggsave(plot = heatmaps,here("fig/fig2_summary_tablev2026.png"), width = 11.82, height = 9.73)
ggsave(plot = heatmaps,here("fig/fig2_summary_tablev2026.pdf"), width = 11.82, height = 9.73)
```

## SI data table

```{r}
supplementary <- as.data.frame(c('ALL',"C only models" ,"CN coupled models","Veg Demo","not Veg Demo",'CLASSIC',
'IBIS',
'LPJ',
'ORCHIDEE',
'VISIT',
'ISBA-CTRIP',
'CABLE-POP',
'CLM5.0',
'DLEM',
'ISAM',
'JSBACH',
'JULES',
'LPJ-GUESS',
'LPX-Bern',
'VISIT-NIES',
'SDGVM'
))
colnames(supplementary) <- c('modl')
supplementary$CN_coupled<-c(NA,NA,NA,NA,NA,'No',
'No',
'No',
'Yes',
'No',
'No',
'Yes',
'Yes',
'Yes',
'Yes',
'Yes',
'Yes',
'Yes',
'Yes',
'No',
'Yes')
supplementary$Veg_Demo<-c(NA,NA,NA,NA,NA,'No',
'No',
'No',
'No',
'No',
'No',
'Yes',
'No',
'No',
'No',
'No',
'No',
'Yes',
'No',
'No',
'Yes')

gdf2<-gdf 

gdf_sampled <- gdf2%>%
  unnest(data) %>%
  bind_rows(
    .,
    gdf %>%
      mutate(n = purrr::map_int(data, ~nrow(.))) %>%
      mutate(data = purrr::map(data, ~slice_sample(., n = nsample,replace=T))) %>% 
      unnest(data) %>%
      mutate(modl = "ALL")
  ) 

Quantile_result <- gdf_sampled%>%
  bind_rows(
    .,
    gdf %>%
      mutate(n = purrr::map_int(data, ~nrow(.))) %>%
      mutate(data = purrr::map(data, ~slice_sample(., n = nsample,replace=T))) %>% # smallest set here
      unnest(data) %>%
      filter(modl %in% c('CLASSIC','IBIS','LPJ', "VISIT" ,  "VISIT-NIES", "ISBA-CTRIP"))%>%
      mutate(modl = "C only models")
  ) %>%
  bind_rows(
    .,
    gdf %>%
      mutate(n = purrr::map_int(data, ~nrow(.))) %>%
      mutate(data = purrr::map(data, ~slice_sample(., n = nsample,replace=T))) %>% # smallest set here
      unnest(data) %>%
      filter(!modl %in% c('CLASSIC','IBIS','LPJ', "VISIT" ,  "VISIT-NIES", "ISBA-CTRIP"))%>%
      mutate(modl = "CN coupled models")
  ) %>%
  bind_rows(
    .,
    gdf %>%
      mutate(n = purrr::map_int(data, ~nrow(.))) %>%
      mutate(data = purrr::map(data, ~slice_sample(., n = nsample,replace=T))) %>% # smallest set here
      unnest(data) %>%
      filter(modl %in% c("CABLE-POP", "LPJ-GUESS", "SDGVM"))%>%
      mutate(modl = "Veg Demo")
  ) %>%
  bind_rows(
    .,
    gdf %>%
      mutate(n = purrr::map_int(data, ~nrow(.))) %>%
      mutate(data = purrr::map(data, ~slice_sample(., n = nsample,replace=T))) %>% # smallest set here
      unnest(data) %>%
      filter(!modl %in% c("CABLE-POP", "LPJ-GUESS", "SDGVM"))%>%
      mutate(modl = "not Veg Demo")
  ) 

# l_cveg_star_npp summary (existing)
Quantile_l_cveg_star_npp <- Quantile_result %>%
  filter(!is.nan(dnpp), !is.nan(dcveg_star), !is.infinite(dnpp), !is.infinite(dcveg_star),
         l_cveg_star_npp < 100000, l_cveg_star_npp > -1000000) %>%
  group_by(modl) %>%
  summarise(
    l_cveg_star_npp_median = quantile(l_cveg_star_npp, 0.5, na.rm=TRUE),
    l_cveg_star_npp_IQR = IQR(l_cveg_star_npp, na.rm=TRUE)
  )


# dcveg → R_cveg
Quantile_R_cveg <- Quantile_result %>%
  filter(!is.nan(dcveg), !is.infinite(dcveg)) %>%
  group_by(modl) %>%
  summarise(
    R_cveg_median = quantile(dcveg, 0.5, na.rm = TRUE),
    R_cveg_IQR = IQR(dcveg, na.rm = TRUE)
  )

# dcveg_ag → R_cveg_ag
Quantile_R_cveg_ag <- Quantile_result %>%
  filter(!is.nan(dcveg_ag), !is.infinite(dcveg_ag)) %>%
  group_by(modl) %>%
  summarise(
    R_cveg_ag_median = quantile(dcveg_ag, 0.5, na.rm = TRUE),
    R_cveg_ag_IQR = IQR(dcveg_ag, na.rm = TRUE)
  )

# dnpp → R_npp
Quantile_R_npp <- Quantile_result %>%
  filter(!is.nan(dnpp), !is.infinite(dnpp)) %>%
  group_by(modl) %>%
  summarise(
    R_npp_median = quantile(dnpp, 0.5, na.rm = TRUE),
    R_npp_IQR = IQR(dnpp, na.rm = TRUE)
  )

# dgpp → R_gpp
Quantile_R_gpp <- Quantile_result %>%
  filter(!is.nan(dgpp), !is.infinite(dgpp)) %>%
  group_by(modl) %>%
  summarise(
    R_gpp_median = quantile(dgpp, 0.5, na.rm = TRUE),
    R_gpp_IQR = IQR(dgpp, na.rm = TRUE)
  )

# dcleaf → R_cleaf
Quantile_R_cleaf <- Quantile_result %>%
  filter(!is.nan(dcleaf), !is.infinite(dcleaf)) %>%
  group_by(modl) %>%
  summarise(
    R_cleaf_median = quantile(dcleaf, 0.5, na.rm = TRUE),
    R_cleaf_IQR = IQR(dcleaf, na.rm = TRUE)
  )

# dcveg_star → R_cveg_star
Quantile_R_cveg_star <- Quantile_result %>%
  filter(!is.nan(dcveg_star), !is.infinite(dcveg_star)) %>%
  group_by(modl) %>%
  summarise(
    R_cveg_star_median = quantile(dcveg_star, 0.5, na.rm = TRUE),
    R_cveg_star_IQR = IQR(dcveg_star, na.rm = TRUE)
  )

# dcroot → R_croot
Quantile_R_croot <- Quantile_result %>%
  filter(!is.nan(dcroot), !is.infinite(dcroot)) %>%
  group_by(modl) %>%
  summarise(
    R_croot_median = quantile(dcroot, 0.5, na.rm = TRUE),
    R_croot_IQR = IQR(dcroot, na.rm = TRUE)
  )

# l_npp_gpp summary
Quantile_l_npp_gpp <- Quantile_result %>%
  filter(!is.nan(l_npp_gpp), !is.infinite(l_npp_gpp)) %>%
  group_by(modl) %>%
  summarise(
    l_npp_gpp_median = quantile(l_npp_gpp, 0.5, na.rm = TRUE),
    l_npp_gpp_IQR = IQR(l_npp_gpp, na.rm = TRUE)
  )

# l_croot_cveg summary
Quantile_l_croot_cveg <- Quantile_result %>%
  filter(!is.nan(l_croot_cveg), !is.infinite(l_croot_cveg)) %>%
  group_by(modl) %>%
  summarise(
    l_croot_cveg_median = quantile(l_croot_cveg, 0.5, na.rm = TRUE),
    l_croot_cveg_IQR = IQR(l_croot_cveg, na.rm = TRUE)
  )

# l_croot_cveg_star summary
Quantile_l_croot_cveg_star <- Quantile_result %>%
  filter(!is.nan(l_croot_cveg_star), !is.infinite(l_croot_cveg_star)) %>%
  group_by(modl) %>%
  summarise(
    l_croot_cveg_star_median = quantile(l_croot_cveg_star, 0.5, na.rm = TRUE),
    l_croot_cveg_star_IQR = IQR(l_croot_cveg_star, na.rm = TRUE)
  )


# l_cleaf_cveg summary
Quantile_l_cleaf_cveg <- Quantile_result %>%
  filter(!is.nan(l_cleaf_cveg), !is.infinite(l_cleaf_cveg)) %>%
  group_by(modl) %>%
  summarise(
    l_cleaf_cveg_median = quantile(l_cleaf_cveg, 0.5, na.rm = TRUE),
    l_cleaf_cveg_IQR = IQR(l_cleaf_cveg, na.rm = TRUE)
  )

# Now combine all summaries together into a single table
Quantile_table <- supplementary %>%
   left_join(Quantile_R_cveg, by = "modl") %>%
  left_join(Quantile_R_cveg_ag, by = "modl") %>%
  left_join(Quantile_R_npp, by = "modl") %>%
  left_join(Quantile_R_gpp, by = "modl") %>%
  left_join(Quantile_R_cleaf, by = "modl") %>%
  left_join(Quantile_R_cveg_star, by = "modl") %>%
  left_join(Quantile_R_croot, by = "modl")%>%
  left_join(Quantile_l_npp_gpp, by = "modl") %>%
  left_join(Quantile_l_croot_cveg, by = "modl") %>%
  left_join(Quantile_l_croot_cveg_star, by = "modl") %>%
  left_join(Quantile_l_cveg_star_npp, by = "modl") %>%
  left_join(Quantile_l_cleaf_cveg, by = "modl")

write.csv(Quantile_table,file = here("data/supplementary_data.csv"))
```

## simple statistic requested by reviewer

```{r}
Quantile_result <- gdf %>%
      mutate(n = purrr::map_int(data, ~nrow(.))) %>%
      mutate(data = purrr::map(data, ~slice_sample(., n = nsample,replace=T))) %>% # smallest set here
      unnest(data) %>%
      filter(modl %in% c('CLASSIC','IBIS','LPJ', "VISIT" ,  "VISIT-NIES", "ISBA-CTRIP"))%>%
      mutate(modl = "C only models") %>%
  bind_rows(
    .,
    gdf %>%
      mutate(n = purrr::map_int(data, ~nrow(.))) %>%
      mutate(data = purrr::map(data, ~slice_sample(., n = nsample,replace=T))) %>% # smallest set here
      unnest(data) %>%
      filter(!modl %in% c('CLASSIC','IBIS','LPJ', "VISIT" ,  "VISIT-NIES", "ISBA-CTRIP"))%>%
      mutate(modl = "CN coupled models")
  ) 

library(dplyr)
library(effectsize)


run_stats <- function(data, value_col, lower = -100, upper = 100) {

  data_clean <- data %>%
    filter(
      !is.na(.data[[value_col]]),
      !is.nan(.data[[value_col]]),
      !is.infinite(.data[[value_col]]),
      .data[[value_col]] > lower,
      .data[[value_col]] < upper
    ) %>%
    mutate(modl = factor(modl, levels = c("C only models", "CN coupled models")))

  group1 <- data_clean %>% filter(modl == "C only models") %>% pull(.data[[value_col]])
  group2 <- data_clean %>% filter(modl == "CN coupled models") %>% pull(.data[[value_col]])

  t_res <- t.test(group1, group2)

  d_res <- effectsize::cohens_d(
    reformulate("modl", response = value_col),
    data = data_clean,
    ci = 0.95
  )

  data.frame(
    Variable = value_col,
    N_C_only_models = length(group1),
    N_CN_coupled_models = length(group2),
    Mean_C_only_models = mean(group1),
    Mean_CN_coupled_models = mean(group2),
    Median_C_only_models = median(group1),
    Median_CN_coupled_models = median(group2),
    T_statistic = unname(t_res$statistic),
    Degrees_of_freedom = unname(t_res$parameter),
    P_value = t_res$p.value,
    Mean_difference_CI_lower = t_res$conf.int[1],
    Mean_difference_CI_upper = t_res$conf.int[2],
    Cohens_d = d_res$Cohens_d,
    Cohens_d_CI_lower = d_res$CI_low,
    Cohens_d_CI_upper = d_res$CI_high
  )
}

df <- Quantile_result %>%
  filter(
    !is.nan(dnpp), !is.nan(dcveg_star),
    !is.infinite(dnpp), !is.infinite(dcveg_star),
    l_cveg_star_npp < 100000, l_cveg_star_npp > -1000000
  ) %>%
  select(modl, l_cveg_star_npp)

df2 <- Quantile_result %>%
  filter(
    !is.nan(l_npp_gpp),
    !is.infinite(l_npp_gpp)
  ) %>%
  select(modl, l_npp_gpp)

df3 <- Quantile_result %>%
  filter(
    !is.nan(l_croot_cveg),
    !is.infinite(l_croot_cveg)
  ) %>%
  select(modl, l_croot_cveg)

res1 <- run_stats(df,  "l_cveg_star_npp", lower = -100, upper = 100)
res2 <- run_stats(df2, "l_npp_gpp",       lower = -100, upper = 100)
res3 <- run_stats(df3, "l_croot_cveg",    lower = -100, upper = 100)

final_table <- bind_rows(res1, res2, res3)

final_table_round <- final_table %>%
  mutate(across(where(is.numeric), ~ round(., 4)))

library(openxlsx)
final_table_t <- as.data.frame(t(final_table_round))
colnames(final_table_t) <- final_table_round$Variable
write.xlsx(
  final_table_t,
  file = "manuscript_statistics_cleaned.xlsx",
  rowNames = TRUE
)



Quantile_result <- gdf %>%
      mutate(n = purrr::map_int(data, ~nrow(.))) %>%
      mutate(data = purrr::map(data, ~slice_sample(., n = nsample,replace=T))) %>% # smallest set here
      unnest(data) %>%
      filter(modl %in% c("CABLE-POP", "LPJ-GUESS", "SDGVM"))%>%
      mutate(modl = "Veg Demo")%>%
  bind_rows(
    .,
    gdf %>%
      mutate(n = purrr::map_int(data, ~nrow(.))) %>%
      mutate(data = purrr::map(data, ~slice_sample(., n = nsample,replace=T))) %>% # smallest set here
      unnest(data) %>%
      filter(!modl %in% c("CABLE-POP", "LPJ-GUESS", "SDGVM"))%>%
      mutate(modl = "not Veg Demo"))



library(dplyr)
library(effectsize)
library(openxlsx)

run_stats <- function(data, value_col, lower = -100, upper = 100) {

  data_clean <- data %>%
    filter(
      !is.na(.data[[value_col]]),
      !is.nan(.data[[value_col]]),
      !is.infinite(.data[[value_col]]),
      .data[[value_col]] > lower,
      .data[[value_col]] < upper
    ) %>%
    mutate(modl = factor(modl, levels = c("Veg Demo", "not Veg Demo")))

  group1 <- data_clean %>% filter(modl == "Veg Demo") %>% pull(.data[[value_col]])
  group2 <- data_clean %>% filter(modl == "not Veg Demo") %>% pull(.data[[value_col]])

  t_res <- t.test(group1, group2)

  d_res <- effectsize::cohens_d(
    reformulate("modl", response = value_col),
    data = data_clean,
    ci = 0.95
  )

  data.frame(
    Variable = value_col,
    N_Veg_Demo = length(group1),
    N_not_Veg_Demo = length(group2),
    Mean_Veg_Demo = mean(group1),
    Mean_not_Veg_Demo = mean(group2),
    Median_Veg_Demo = median(group1),
    Median_not_Veg_Demo = median(group2),
    T_statistic = unname(t_res$statistic),
    Degrees_of_freedom = unname(t_res$parameter),
    P_value = t_res$p.value,
    Mean_difference_CI_lower = t_res$conf.int[1],
    Mean_difference_CI_upper = t_res$conf.int[2],
    Cohens_d = d_res$Cohens_d,
    Cohens_d_CI_lower = d_res$CI_low,
    Cohens_d_CI_upper = d_res$CI_high
  )
}

df <- Quantile_result %>%
  filter(
    !is.nan(dnpp), !is.nan(dcveg_star),
    !is.infinite(dnpp), !is.infinite(dcveg_star),
    l_cveg_star_npp < 100000,
    l_cveg_star_npp > -1000000
  ) %>%
  select(modl, l_cveg_star_npp)

df2 <- Quantile_result %>%
  filter(
    !is.nan(l_npp_gpp),
    !is.infinite(l_npp_gpp)
  ) %>%
  select(modl, l_npp_gpp)

df3 <- Quantile_result %>%
  filter(
    !is.nan(l_croot_cveg),
    !is.infinite(l_croot_cveg)
  ) %>%
  select(modl, l_croot_cveg)

res1 <- run_stats(df,  "l_cveg_star_npp", lower = -100, upper = 100)
res2 <- run_stats(df2, "l_npp_gpp",       lower = -100, upper = 100)
res3 <- run_stats(df3, "l_croot_cveg",    lower = -100, upper = 100)

final_table <- bind_rows(res1, res2, res3)

final_table_round <- final_table %>%
  mutate(across(where(is.numeric), ~ round(., 4)))

print(final_table_round)

# Transpose for manuscript-friendly layout
final_table_t <- as.data.frame(t(final_table_round))
colnames(final_table_t) <- final_table_round$Variable

# Save to Excel
write.xlsx(
  final_table_t,
  file = "manuscript_statistics_veg_demo.xlsx",
  rowNames = TRUE
)

```


